### 2010-12-06 cs886-sarah/Paper.tex

\documentclass{article}

% For figures
\usepackage{graphicx}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2010, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2010} with
% \usepackage[nohyperref]{icml2010} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage[accepted]{icml2010}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Title}

\begin{document}

\twocolumn[
\icmltitle{Using Bayesian networks to predict changes}

\icmlauthor{Sarah Nadi}{snadi@uwaterloo.ca}
\icmladdress{University of Waterloo,
            Waterloo, ON, Canada}

\vskip 0.3in
]

\begin{abstract} 
STILLLL
\end{abstract}

\section{Introduction}
\label{intro}

Changes

\section{Background}

\subsection{CMDBs and Change Sets}
A Configuration Management Database (CMDB) is useful in Enterprise IT Management (EITM) since it provides information about the various critical components in a
system including hardware, software, and services provided by the company. It records the configuration of these items, their change history, their incident
history, as well as the relationships between them. Each item stored in the CMDB is referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows
an example CMDB to illustrate the concepts of CIs and relationships. In this example, there are two software applications being hosted by two different web
servers. Both applications communicate with the same databases through a load balancer. Such a visualization allows an IT analyst to better understand the
system at hand.

A CMDB provides a basis for decision making processes such as Incident Management, Change Management, etc.  In this paper, we focus on the process of Change
Management, and in particular, on the problem of change set detection .A \textit{change} is the addition, modification, or removal of anything that could affect
on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change one CI is the system, other CIs that might
need to be changed as well must be correctly identified. The set of CIs that will need to be modified for the change to be complete is called a change set.

\begin{figure}[!t]
\centering
\includegraphics[width=7cm]{cmdbExample.PNG}
\caption{An IT System Stored in a CMDB}
\label{fig:cmdbExample}
\end{figure}

\subsection{Bayesian Networks}

\subsubsection{Tools}

\subsubsection*{Weka}

\subsubsection*{Banjo}
Banjo is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms: Greedy and
Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
which basically proposes a random addition or removal of an edge. The other is ``proposeAllLocalMoves'' which proposes all possible moves, and only keeps the
best one (?). Banjo uses the BDe metric to compute a network's score.

\subsubsection*{JavaBayes}

\subsection{Evaluation Techniques}

We need a way to evaluate the predicted CIs. The recall and precision measures from the information retrieval field are appropriate for this type of evaluation.
Recall measures the proportion of correct CIs retrieved by the system, while precision measures the proportion of suggested CIs that are correct~\cite{van79}.

Similar to Hassan et. al~\cite{hassan2004predicting}, we define the \textit{Predicted Set} (P) as the set of all CIs DRACA suggests through the full the
iteration process (see Figure~\ref{fig:process}). We define the \textit{Occurred Set} (O) as the CIs remaining in the change set after excluding the Initial CI
provided by the analyst (i.e Change Set - Initial CI). The intersection of the predicted set and the occurred set, called $PO$, is the common CIs in both sets.
For each constructed change set, we then calculate the recall and precision values for the predictions according to the following
definitions~\cite{hassan2004predicting}:

\begin{equation}
\label{eqn:recall}
Recall = \frac{|PO|}{|O|}
\end{equation}

\begin{equation}
\label{eqn:precision}
Precision = \frac{|PO|}{|P|}
\end{equation}

If no CIs are predicted (i.e., $P$ and thus $PO$ are empty), precision is defined as 1 since there cannot exist any incorrect predictions in an empty set. On
the other hand, if the size of the change set is 1, and thus the size of the occurred set is 0, recall is defined as 1 since there are no CIs to
predict~\cite{hassan2004predicting}.

In order to have a single measure that indicates the effectiveness of our predictions, we use the F-measure which is based on van Rijsbergen's effectiveness
measure which combines recall and precision~\cite{van79}. The F-measure is calculated according to Equation~\ref{eqn:f-measure} which gives equal weighting to
recall and precision. The ideal F-measure is 1 where both recall and precision are 1.

\begin{equation}
F = 2 * \frac{precision * recall}{precision +recall}
\label{eqn:f-measure}
\end{equation}

\section{Related Work}
\label{rel-work}

Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. However, their work is on the level of source code changes. They build three
different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies, and then the CPTs
are learnt using the importance sampling algorithm proposed by Changhe and Marek~\cite{yuan2003importance}. The way static dependencies are defined in their
case is specific to Java. The third one is essentially the first graph, but updated using the historic change information according to the Expectation
Maximization (EM) algorithm~\cite{dempster1977maximum}. The second was solely based on historic information where the network is build using a greedy structure
 learning algorithm~\cite{friedman1996learning}. They did some preprocessing to their data such as filtering out large changes (with more than 30 elements
changed at once) since this was probably an insignificant change.

Zhou et al.~\cite{zhou2008} try to answer a slightly different problem. They do not only look at the probability of other elements changing given a specific
element, they also add features such as authors, change significance levels etc. and try to predict if two elements are co-changes or not accordingly. Thus,
their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as co-changes or not.
They use the K2 algorithm proposed by Cooper et. al~\cite{cooper1992bayesian} to estimate the structure of the Bayesian network, and use the SimpleEstimator
algorithm built in WEKA~\cite{witten2005data}.

\section{Models Used}
\label{sec:modelsused}

We experimented with several models to predict change sets. First, we just used Weka's BayesNet classifier to build the structure and estimate the CPTs of
the network. In the second model, we used Banjo to learn the structure of the network, and then we used Weka to learn the CPTs. In the third model, we imposed
the structure of the network from the relationships already existing between CIs in the CMDB. Finally, in the fourth model, we tried to assume that CIs not in
a change order are unobserved data, and therefore have missing values.

In order to be able to test things properly, we chose a small dataset to be able to build our model. We used observations from three months data from January 1,
2008 to March 31, 2008 to build the model. However, in such a short period, there were already 2,841 distinct CIs appearing and 2,229 observations (i.e. change
orders). Therefore, in each model, some data preprocessing was done before building the model since having a network with over 2000 variables was very hard to
deal with.

\subsection{Model 1}
\label{sec:model1}

\subsection{Model 2}
\label{sec:model2}

\subsubsection*{Data Preprocessing}

For this model, we first got a list of all the CIs that appeared more than 12 times (i.e were associated with 12 different change orders in our training
observations). This yielded 120 distinct CIs. Then, in order to slightly increase our variable space to include other related CIs, we found all the parent CIs
related to these 120 CIs from the CMDB perspective. However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is
location for'', and ``backs up''. Adding the related parent CIs, we now had a set of 241 CIs. We then went back, and iterated through the observations, and
kept only observations for CIs that either had more than 12 different change orders or were related to these CIs (i.e that occurred in the set of 241 CIs).
This, finally, gave us a set of 170 distinct CIs which we used for our model. Additionally, filtering out CIs meant filtering out some of the observations that
did not have any CIs satisfying our criteria. This led to us having 1305 observations instead of 2841 which was a more manegeable set.

\subsubsection*{Building the Model}

Given that we had our training set, the next step was to build the network which has two parts: building the structure, and estimating the CPTs. Unfortunately,
Weka mainly focused on classification and clustering which is not exactly the same problem we are tackling here. However, Weka did provide a functionality to
learn the CPT of a network given a certain structure. We, therefore, used Banjo to build the structure of the network (Banjo does not estimate CPTs), and then
used Weka to estimate the CPTs on the produced network. In both cases, the same observations file was used to ensure that its the same data. However, each tool
accepted the data in different formats, so some translation was necessary there. After Weka has added the CPT information to the graph, we used this as the
final Bayesian network on which to do our analysis.

\subsubsection*{Inference}

At this point, we have the Bayesian network ready, and we would like to perform Inference. More formally, given that a CI will change (our evidence), we want
to infer the probability that the other CIs in the network might change as well. Unfortunately, neither of the two tools previously used provide an Bayesian
inference engine. We, therefore, use JavaBayes in this step since it accepts the same ARFF format used by Weka. We using a one month testing set where we try
to predict all the change orders in April 2008 (883 reports). There was a total of XX change orders in that month. For each change order, we would take the
first CI as the initial CI to change, then we would set that as evidence, and update the beliefs of all the nodes (CIs) in the network using JavaBayes. We would
then loop on all the updated CIs, and add those that match our threshold criteria to the predicted change set. To simulate a real life scenario, we then checked
which of these predicted CIs actually lies in the target change set we are trying to predict. This is similar to an analyst accepting CIs into their change set.
These common CIs would then also be marked as observations so that we can predict now that I'm going to change B and C too, what else do I need to change.
Again, we would calculate the posterior probability, and continue doing so until there are no more common CIs. We would then calculate the recall, precision,
and F-measure accordingly.

\subsubsection*{Results}

\subsection{Model 3}

\subsubsection*{Data Preprocessing}

For this model, we only used the list of CIs appearing more than 12 times (120 distinct CIs).

\subsubsection*{Building the Model}

Using CIs in this list only, we constructed the structure of the Bayesian network by adding in CMDB relationships between the CIs. We stuck to this set only to
ensure that all the CIs in the network, are also present in the observations. Fourty-eight edges were added to the structure of the network. Using these 120
CIs, we had XXX observations. Again, we loaded the produced structure in Weka, and learned the CPTs using the observations from the three months. We then took
the produced Bayesian network, and used JavaBayes to perform the inference.

\subsubsection*{Inference}

\section{Building the model}
\label{sec:model}

\section{Experimental Work}
\label{sec:exp}

\section{Conclusion}
\label{concl}

\bibliography{references}
\bibliographystyle{icml2010}

\end{document}

### 2010-12-07 trunk/cs886-sarah/Paper.tex
added figure and algorithm

--- 2010-12-06 cs886-sarah/Paper.tex
@@
 
-We experimented with several models to predict change sets. First, we just used Weka's BayesNet classifier to build the structure and estimate the CPTs of
-the network. In the second model, we used Banjo to learn the structure of the network, and then we used Weka to learn the CPTs. In the third model, we imposed
-the structure of the network from the relationships already existing between CIs in the CMDB. Finally, in the fourth model, we tried to assume that CIs not in
-a change order are unobserved data, and therefore have missing values.
+In order to construct a Bayesian network which we can use to make predictions, there were two steps involved. First, determining the structure of the actual
+network, and then estimating the Conditional Probability Tables (CPT). There were different ways in which we could estimate the structure of the network, and
+so we built a model for each technique. However, for all models, we used the SimpleEstimator algorithm~\cite{written2005data} built in Weka to calculate the
+CPTs.
 
-In order to be able to test things properly, we chose a small dataset to be able to build our model. We used observations from three months data from January 1,
-2008 to March 31, 2008 to build the model. However, in such a short period, there were already 2,841 distinct CIs appearing and 2,229 observations (i.e. change
-orders). Therefore, in each model, some data preprocessing was done before building the model since having a network with over 2000 variables was very hard to
-deal with.
+Before building any of the models, some data preprocessing was necessary. We first describe this preprocessing, and then proceed to explain each of the FOUR
+models built.
 
-\subsection{Model 1}
+\begin{figure}[!t]
+\centering
+\includegraphics[width=7cm]{graphics/constructingmodel.png}
+\caption{Generating the Bayesian Network}
+\label{fig:process}
+\end{figure}
+
+\subsubsection{Data Preprocessing}
+The original data set has 7,999 distinct CIs, and 27,305 change orders. This amount of data was infeasible to work with as no tool could handle such a large
+amount of variables in a Bayesian network. For the purposes of this project, which is mainly to experiment with Bayesian techniques for change set prediction,
+it is sufficient to choose a small representative subset of the data. Accordingly, in order to be able to test things properly, we used observations from three
+months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a short period, there were already 2,841 distinct CIs appearing and
+2,229 observations (i.e. change orders). Therefore, we needed to perform further data preprocessing.
+
+First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with 12 different change orders in our training
+observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct CIs. Then,
+in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB perspective.
+However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is location for'', and ``backs up''. Adding the related
+parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs. Accordingly, we just kept CIs tht
+appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use throughout our models for fair
+comparison. Additionally, filtering out CIs meant filtering out some of the observations that did not have any CIs satisfying our criteria. This led to us
+having 1,305 observations instead of 2,229 which was a more manegeable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305
+observations (change sets).
+
+DESCRIBE GENERATED OBSERVATIONS FILE
+
+FIGURE FOR DATA PROCESSING
+
+\subsection{Model 1 (Banjo)}
 \label{sec:model1}
 
-\subsection{Model 2}
-\label{sec:model2}
+For the first model, we used Banjo to learn the structure of the Bayesian network from the training data which used a Greedy searching algorithm. Banjo
+produced five different top-scoring networks (having the same score), and we simply chose one of them as the network to be used. This network was then
+formatted as an BIF XML file, and was fed into Weka's network analyzer tool. After setting the data set to be the observations in our training set, Weka
+learned the CPTs for the network produced by Banjo. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference.
 
-\subsubsection*{Data Preprocessing}
+\subsection{Model 2 (CMDB Relationships)}
 
-For this model, we first got a list of all the CIs that appeared more than 12 times (i.e were associated with 12 different change orders in our training
-observations). This yielded 120 distinct CIs. Then, in order to slightly increase our variable space to include other related CIs, we found all the parent CIs
-related to these 120 CIs from the CMDB perspective. However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is
-location for'', and ``backs up''. Adding the related parent CIs, we now had a set of 241 CIs. We then went back, and iterated through the observations, and
-kept only observations for CIs that either had more than 12 different change orders or were related to these CIs (i.e that occurred in the set of 241 CIs).
-This, finally, gave us a set of 170 distinct CIs which we used for our model. Additionally, filtering out CIs meant filtering out some of the observations that
-did not have any CIs satisfying our criteria. This led to us having 1305 observations instead of 2841 which was a more manegeable set.
+For the second model, we used the relations existing in the CMDB to infer the structure of the network (that is place the edges between the nodes). We used two
+variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
+edge in the Bayesian network from A to B (we will call this model 2A). The second was to reverse the direction of the relationship edges in the CMDB. That is,
+if there is an edge from A to B, we will place an edge in the Bayesian network from B to A (we will call this model 2B). In both cases, however, we chose to
+ignore some relationship edges (WHY, TRY WITHOUT). There was one problem, however, with the generated network. Two of the CIs had more than 15 parents to them
+which means that their CPTs will be intractable to compute. For those two CIs, we simply removed all their parents to make the computation tractable.
 
-\subsubsection*{Building the Model}
+\subsection{Model 3 (Missing Data)}
 
-Given that we had our training set, the next step was to build the network which has two parts: building the structure, and estimating the CPTs. Unfortunately,
-Weka mainly focused on classification and clustering which is not exactly the same problem we are tackling here. However, Weka did provide a functionality to
-learn the CPT of a network given a certain structure. We, therefore, used Banjo to build the structure of the network (Banjo does not estimate CPTs), and then
-used Weka to estimate the CPTs on the produced network. In both cases, the same observations file was used to ensure that its the same data. However, each tool
-accepted the data in different formats, so some translation was necessary there. After Weka has added the CPT information to the graph, we used this as the
-final Bayesian network on which to do our analysis.
+This model is the same as the second model, except that we experimented with the missing data feature in Weka. Weka allows the user to specify missing values
+for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
+'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. Therefore, this model also has
+two parts: model 3A and model 3B where the first uses the CMDB relationships in their same direction, and the second uses the reverse direction.
+
+\begin{algorithm}[tb]
+   \caption{Bubble Sort}
+   \label{alg:example}
+\begin{algorithmic}
+   \STATE {\bfseries Input:} changeOrder
+    \STATE {\bfseries Input:} BayesianNetwork
+    \STATE {\bfseries Input:} threshold
+    \STATE Set initialCI = first CI in changeOrder
+    \STATE Set occurredSet = changeOrder - initialCI
+    \STATE Initialize observedSet = {firstCI}
+    \STATE Initialize predictedSet = {}
+   \REPEAT  
+    \STATE Initialize newPredictions = {}
+    \FOR{$node$ {\bfseries to} $BayesianNetwork$}
+    \STATE posterior = perform inference using Banjo
+    \IF{$posterior > threshold$} 
+      \STATE Add node to newPredictions   
+    \ENDIF
+    \ENDFOR
+
+  \FOR{$prediction$ {\bfseries to} $newPredictions$}
+    \IF{$prediction  occurredSet$}
+      \STATE Add prediction to observedSet
+    \ENDIF
+    \ENDFOR
+\STATE Add newPredictions to predictedSet
+
+   \UNTIL{$newPredictions$ is $empty$}
+\end{algorithmic}
+\end{algorithm}
+
+\section{Experimental Work}
+\label{sec:exp}
 
@@
 
-\subsubsection*{Results}
-
-\subsection{Model 3}
-
-\subsubsection*{Data Preprocessing}
-
-For this model, we only used the list of CIs appearing more than 12 times (120 distinct CIs).
-
-\subsubsection*{Building the Model}
-
-Using CIs in this list only, we constructed the structure of the Bayesian network by adding in CMDB relationships between the CIs. We stuck to this set only to
-ensure that all the CIs in the network, are also present in the observations. Fourty-eight edges were added to the structure of the network. Using these 120
-CIs, we had XXX observations. Again, we loaded the produced structure in Weka, and learned the CPTs using the observations from the three months. We then took
-the produced Bayesian network, and used JavaBayes to perform the inference.
-
-\subsubsection*{Inference}
-
-\section{Building the model}
-\label{sec:model}
-
-\section{Experimental Work}
-\label{sec:exp}
-
 \section{Conclusion}

### 2010-12-07 cs886-sarah/Paper.tex

@@
 \usepackage{algorithmic}
+\usepackage{url}
 
@@
 
-Changes
+Change is inevitable in any Information Technology(IT) system. New features are added, different configurations are used, upgrades are introduced and new
+software and hardware are added. In a large system, such changes can cause other parts of the system to malfunction without the ability of the analyst to
+foresee this side effect. Accordingly, we need to ensure that we consider the impact of all changes before they are implemented, and make sure we adjust the
+potentially affected components accordingly while implementing the change. This leads to the process of change set detection which is finding all the
+components of the system that need to be included in your change set where the change set is the set of components that need to be changed to avoid any side
+effects.
+
+Many of the large organizations with large and sophisticated IT systems employ a Configuration Management Database (CMDB) to help them manage these systems.
+The CMDB keeps track of all the components in a system, how they are related, as well as all the changes that have occurred to them. Any component in the CMDB
+is called a Configuration Item (CI). In previous work~\cite{nadi2010}, we mined the CMDB repository for historical co-changes, and used the mined correlations
+to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well based on its change
+history with other CIs. We obtained really good results in terms of recall and precision (69.8\% and 88.5\% respectively).
+
+In this paper, we wish to examine the same problem, but from a Bayesian perspective. Instead of looking at pairwise historical co-changes of CI, we would like
+to consider all the changed CIs in the past to deduce the causality relations between them. That is, what is the probability that if A changes, B changes as
+well. In that case, A would be a parent to B in a Bayesian network. Accordingly, we explore the different ways a Bayesian network can be constructed from the
+data we have, and then test the predictions produced by querying this network to examine the obtained recall and precision. We hope that we can obtain better
+results using Bayesian networks since they are not limited to pairwise comparisons.
+
+The rest of this paper is organized as follows. BLA BLA BLA
 
@@
 \subsection{CMDBs and Change Sets}
-A Configuration Management Database (CMDB) is useful in Enterprise IT Management (EITM) since it provides information about the various critical components in a
+A Configuration Management Database (CMDB)~\cite{itil} is useful in Enterprise IT Management (EITM) since it provides information about the various critical
+components in a
 system including hardware, software, and services provided by the company. It records the configuration of these items, their change history, their incident
 history, as well as the relationships between them. Each item stored in the CMDB is referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows
-an example CMDB to illustrate the concepts of CIs and relationships. In this example, there are two software applications being hosted by two different web
-servers. Both applications communicate with the same databases through a load balancer. Such a visualization allows an IT analyst to better understand the
-system at hand.
+an example CMDB to illustrate the concepts of CIs and relationships.
 
 A CMDB provides a basis for decision making processes such as Incident Management, Change Management, etc.  In this paper, we focus on the process of Change
-Management, and in particular, on the problem of change set detection .A \textit{change} is the addition, modification, or removal of anything that could affect
+Management, and in particular, on the problem of change set detection. A \textit{change} is the addition, modification, or removal of anything that could affect
 on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change one CI is the system, other CIs that might
@@
 \includegraphics[width=7cm]{cmdbExample.PNG}
-\caption{An IT System Stored in a CMDB}
+\caption{An IT System Stored in a CMDB. Each item shown is a CI, and CIs are tied through different relationship types}
 \label{fig:cmdbExample}
@@
 
-\subsubsection{Tools}
+\subsection{Bayesian Network Tools}
 
-\subsubsection*{Weka}
+\subsubsection{Weka}
 
-\subsubsection*{Banjo}
-Banjo is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms: Greedy and
-Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
+WEKA~\cite{hall2009weka} is a toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
+learning problems. It also provides Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
+Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. We will mainly be using WEKA for learning the CPTs in a fixed
+structure network using the SimpleEstimator algorithm~\cite{written2005data} implemented there. This simple estimates the probability values on edges based on
+the frequency values of each variable given its parents in the training set.
+
+\subsubsection{Banjo}
+Banjo~\cite{banjotool} is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms:
+Greedy and Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
 which basically proposes a random addition or removal of an edge. The other is ``proposeAllLocalMoves'' which proposes all possible moves, and only keeps the
@@
 
-\subsubsection*{JavaBayes}
+\subsubsection{JavaBayes}
+
+\section{Related Work}
+\label{rel-work}
+
+Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. However, their work is on the level of source code changes. They build three
+different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
+co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies, and then the CPTs
+are learnt using the importance sampling algorithm proposed by Changhe and Marek~\cite{yuan2003importance}. The way static dependencies are defined in their
+case is specific to Java. The third one is essentially the first graph, but updated using the historic change information according to the Expectation
+Maximization (EM) algorithm~\cite{dempster1977maximum}. The second was solely based on historic information where the network is build using a greedy structure
+ learning algorithm~\cite{friedman1996learning}. They did some preprocessing to their data such as filtering out large changes (with more than 30 elements
+changed at once) since this was probably an insignificant change.
+
+Zhou et al.~\cite{zhou2008} try to answer a slightly different problem. They do not only look at the probability of other elements changing given a specific
+element, they also add features such as authors, change significance levels etc. and try to predict if two elements are co-changes or not accordingly. Thus,
+their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as co-changes or not.
+They use the K2 algorithm proposed by Cooper et. al~\cite{cooper1992bayesian} to estimate the structure of the Bayesian network, and use the SimpleEstimator
+algorithm built in WEKA~\cite{witten2005data}.
+
+\section{Constructing the Bayesian Network}
+\label{sec:modelsused}
+
+Before constructing the Bayesian Network, we first had to process and prepare the data we have. Then, in order to construct a Bayesian network to use for
+predictions there are two steps involved. First, determining the structure of the actual network, and then estimating the Conditional Probability Tables
+(CPT). Section~\ref{sec:data} first explains the data set available to us, and the data preprocessing involved before building the network.
+Section~\ref{sec:structure} then explains the different techniques we experimented with to build the network structure. Section~\ref{sec:cpt} shows the last
+step to build the network which is estimating the CPTs. Figure~\ref{fig:process} shows this overall process.
+
+\subsection{Data Preprocessing}
+\label{sec:data}
+
+\subsubsection*{Data Size}
+The original data set is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data was infeasible to work with as no tool
+could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to experiment with Bayesian techniques
+for change set prediction,
+it is sufficient to choose a small representative subset of the data. Accordingly, in order to be able to test things properly, we used observations from three
+months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a short period, there were already 2,841 distinct CIs appearing and
+2,229 observations (i.e. change orders). Therefore, we needed to perform further data preprocessing.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=7cm]{graphics/constructingmodel.png}
+\caption{Building the Bayesian Network}
+\label{fig:process}
+\end{figure}
+
+First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with 12 different change orders in our training
+observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct CIs. Then,
+in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB perspective.
+However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is location for'', and ``backs up''. Adding the related
+parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs. Accordingly, we just kept CIs tht
+appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use throughout our models for fair
+comparison. Additionally, filtering out CIs meant filtering out some of the observations that did not have any CIs satisfying our criteria. This led to us
+having 1,305 observations instead of 2,229 which was a more manegeable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305
+observations (change sets).
+
+FIGURE FOR DATA PROCESSING
+
+\subsubsection*{Data Format}
+\label{sec:format}
+
+In the CMDB, a Change Order has several fields including the requester, the assignee, the start date of the change, the description and the change set field
+(called the `Configuration items' fields). Of the fields in a change order, we use only the change set field. The advantage of only using this one field is that
+change sets are easy to extract and easy to understand. In terms of an observation, CIs were either marked as ``true'' or ``false'' indicating whether they
+have changed or not. Each observation initially consisted of all the in our data set marked as ``false''. For each change order in the training set, we would
+mark each CI appearing in the ``Configuration Items'' field as true, while all those not appearing were left as ``false''. Therefore, the set of training
+data consisted of an observation entry for each change order with the CIs marked as ``true'' or ``false'' accordingly.
+
+\subsection{Network Structure}
+\label{sec:structure}
+There were different ways in which we could estimate the structure of the network, and so we built a Bayesian Network using each technique.
+
+\subsubsection{Model 1 (Banjo)}
+\label{sec:model1}
+
+For the first model, we used Banjo to learn the structure of the Bayesian network from the training data. We use the Greedy searching algorithm implemented
+there. Banjo produced five different top-scoring networks (having the same score), and we simply chose one of them as the network to be used.
+
+\subsubsection{Model 2 (CMDB Relationships)}
+
+For the second model, we used the relations existing in the CMDB to infer the structure of the network (that is place the edges between the nodes). We used two
+variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
+edge in the Bayesian network from A to B (we will call this model 2A). The second was to reverse the direction of the relationship edges in the CMDB. That is,
+if there is an edge from A to B, we will place an edge in the Bayesian network from B to A (we will call this model 2B). In both cases, however, we chose to
+ignore some relationship edges (WHY, TRY WITHOUT). There was one problem, however, with the generated network. Two of the CIs had more than 15 parents to them
+which means that their CPTs will be intractable to compute. For those two CIs, we simply removed all their parents to make the computation tractable.
+
+\subsubsection{Model 3 (Missing Data)}
+
+This model is the same as the second model, except that we experimented with the missing data feature in Weka. Weka allows the user to specify missing values
+for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
+'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. Therefore, this model also has
+two parts: model 3A and model 3B where the first uses the CMDB relationships in their same direction, and the second uses the reverse direction.
+
+\subsection{Estimating the CPTs}
+\label{sec:cpt}
+
+However, for all
+models, we used the SimpleEstimator algorithm~\cite{written2005data} built in Weka to calculate the
+CPTs. After setting the data set to be the observations in our training set, Weka
+learned the CPTs for the network produced by Banjo. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference.
+
+\section{Experiment Setup}
+
+\begin{algorithm}[tb]
+   \caption{Generating Predictions using the Bayesian Network}
+   \label{alg:simulation}
+\begin{algorithmic}
+   \STATE {\bfseries Input:} changeOrder
+    \STATE {\bfseries Input:} BayesianNetwork
+    \STATE {\bfseries Input:} threshold
+    \STATE Set initialCI = first CI in changeOrder
+    \STATE Set occurredSet = changeOrder - initialCI
+    \STATE Initialize observedSet = {firstCI}
+    \STATE Initialize predictedSet = {}
+   \REPEAT  
+    \STATE Initialize newPredictions = {}
+    \FOR{$node$ {\bfseries to} $BayesianNetwork$}
+    \STATE posterior = perform inference using Banjo
+    \IF{$posterior > threshold$} 
+      \STATE Add node to newPredictions   
+    \ENDIF
+    \ENDFOR
+
+  \FOR{$prediction$ {\bfseries to} $newPredictions$}
+    \IF{$prediction  occurredSet$}
+      \STATE Add prediction to observedSet
+    \ENDIF
+    \ENDFOR
+\STATE Add newPredictions to predictedSet
+
+   \UNTIL{$newPredictions$ is $empty$}
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Change Set Detection Process Simulation}
+
+The change set detection process provided is an iterative, collaborative process between the tool and the analyst. When the tool suggests CIs, the analyst can
+accept or reject these CIs (i.e add them to the change set or not). Based on the CIs added to the change set, the analyst can ask the tool for more
+suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''.
+
+To simulate this process for our experiments, we do the following. At this point, we have the Bayesian network ready, and we would like to perform Inference.
+More formally, given that a CI will change (our observation), we want to infer the probability that the other CIs in the network might change as well.
+Unfortunately, neither of the two tools previously used provide an Bayesian inference engine. We, therefore, use JavaBayes in this step since it accepts the
+same ARFF format used by Weka. We using a one month testing set where we try to predict all the change orders in April 2008. There was a total of
+883 change orders in that month.
+
+For each change order, we would take the first CI as the initial CI to change, then we would set that as an observed node, and update
+the beliefs of all the nodes (CIs) in the network using JavaBayes. We would then loop on all the updated CIs, and add those that match our threshold criteria to
+the predicted change set. To simulate a real life scenario, we then checked which of these predicted CIs actually lies in the target change set we are trying to
+predict. This is similar to an analyst accepting CIs into their change set. These common CIs would then also be marked as observations so that we can predict
+now that I'm going to change these accepted CIs as well, what else do I need to change. Again, we would calculate the posterior probability, and continue doing
+so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst or not) are part of the predicted
+set.
 
@@
 
-\section{Related Work}
-\label{rel-work}
+Algorithm~\ref{alg:simulation} shows this process, and how the recall and precision are calculated accordingly.
 
-Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. However, their work is on the level of source code changes. They build three
-different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
-co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies, and then the CPTs
-are learnt using the importance sampling algorithm proposed by Changhe and Marek~\cite{yuan2003importance}. The way static dependencies are defined in their
-case is specific to Java. The third one is essentially the first graph, but updated using the historic change information according to the Expectation
-Maximization (EM) algorithm~\cite{dempster1977maximum}. The second was solely based on historic information where the network is build using a greedy structure
- learning algorithm~\cite{friedman1996learning}. They did some preprocessing to their data such as filtering out large changes (with more than 30 elements
-changed at once) since this was probably an insignificant change.
-
-Zhou et al.~\cite{zhou2008} try to answer a slightly different problem. They do not only look at the probability of other elements changing given a specific
-element, they also add features such as authors, change significance levels etc. and try to predict if two elements are co-changes or not accordingly. Thus,
-their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as co-changes or not.
-They use the K2 algorithm proposed by Cooper et. al~\cite{cooper1992bayesian} to estimate the structure of the Bayesian network, and use the SimpleEstimator
-algorithm built in WEKA~\cite{witten2005data}.
-
-\section{Models Used}
-\label{sec:modelsused}
-
-We experimented with several models to predict change sets. First, we just used Weka's BayesNet classifier to build the structure and estimate the CPTs of
-the network. In the second model, we used Banjo to learn the structure of the network, and then we used Weka to learn the CPTs. In the third model, we imposed
-the structure of the network from the relationships already existing between CIs in the CMDB. Finally, in the fourth model, we tried to assume that CIs not in
-a change order are unobserved data, and therefore have missing values.
-
-In order to be able to test things properly, we chose a small dataset to be able to build our model. We used observations from three months data from January 1,
-2008 to March 31, 2008 to build the model. However, in such a short period, there were already 2,841 distinct CIs appearing and 2,229 observations (i.e. change
-orders). Therefore, in each model, some data preprocessing was done before building the model since having a network with over 2000 variables was very hard to
-deal with.
-
-\subsection{Model 1}
-\label{sec:model1}
-
-\subsection{Model 2}
-\label{sec:model2}
-
-\subsubsection*{Data Preprocessing}
-
-For this model, we first got a list of all the CIs that appeared more than 12 times (i.e were associated with 12 different change orders in our training
-observations). This yielded 120 distinct CIs. Then, in order to slightly increase our variable space to include other related CIs, we found all the parent CIs
-related to these 120 CIs from the CMDB perspective. However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is
-location for'', and ``backs up''. Adding the related parent CIs, we now had a set of 241 CIs. We then went back, and iterated through the observations, and
-kept only observations for CIs that either had more than 12 different change orders or were related to these CIs (i.e that occurred in the set of 241 CIs).
-This, finally, gave us a set of 170 distinct CIs which we used for our model. Additionally, filtering out CIs meant filtering out some of the observations that
-did not have any CIs satisfying our criteria. This led to us having 1305 observations instead of 2841 which was a more manegeable set.
-
-\subsubsection*{Building the Model}
-
-Given that we had our training set, the next step was to build the network which has two parts: building the structure, and estimating the CPTs. Unfortunately,
-Weka mainly focused on classification and clustering which is not exactly the same problem we are tackling here. However, Weka did provide a functionality to
-learn the CPT of a network given a certain structure. We, therefore, used Banjo to build the structure of the network (Banjo does not estimate CPTs), and then
-used Weka to estimate the CPTs on the produced network. In both cases, the same observations file was used to ensure that its the same data. However, each tool
-accepted the data in different formats, so some translation was necessary there. After Weka has added the CPT information to the graph, we used this as the
-final Bayesian network on which to do our analysis.
-
-\subsubsection*{Inference}
-
-At this point, we have the Bayesian network ready, and we would like to perform Inference. More formally, given that a CI will change (our evidence), we want
-to infer the probability that the other CIs in the network might change as well. Unfortunately, neither of the two tools previously used provide an Bayesian
-inference engine. We, therefore, use JavaBayes in this step since it accepts the same ARFF format used by Weka. We using a one month testing set where we try
-to predict all the change orders in April 2008 (883 reports). There was a total of XX change orders in that month. For each change order, we would take the
-first CI as the initial CI to change, then we would set that as evidence, and update the beliefs of all the nodes (CIs) in the network using JavaBayes. We would
-then loop on all the updated CIs, and add those that match our threshold criteria to the predicted change set. To simulate a real life scenario, we then checked
-which of these predicted CIs actually lies in the target change set we are trying to predict. This is similar to an analyst accepting CIs into their change set.
-These common CIs would then also be marked as observations so that we can predict now that I'm going to change B and C too, what else do I need to change.
-Again, we would calculate the posterior probability, and continue doing so until there are no more common CIs. We would then calculate the recall, precision,
-and F-measure accordingly.
-
-\subsubsection*{Results}
-
-\subsection{Model 3}
-
-\subsubsection*{Data Preprocessing}
-
-For this model, we only used the list of CIs appearing more than 12 times (120 distinct CIs).
-
-\subsubsection*{Building the Model}
-
-Using CIs in this list only, we constructed the structure of the Bayesian network by adding in CMDB relationships between the CIs. We stuck to this set only to
-ensure that all the CIs in the network, are also present in the observations. Fourty-eight edges were added to the structure of the network. Using these 120
-CIs, we had XXX observations. Again, we loaded the produced structure in Weka, and learned the CPTs using the observations from the three months. We then took
-the produced Bayesian network, and used JavaBayes to perform the inference.
-
-\subsubsection*{Inference}
-
-\section{Building the model}
-\label{sec:model}
-
-\section{Experimental Work}
-\label{sec:exp}
+\section{Results}
+\label{sec:results}
 

### 2010-12-08 trunk/cs886-sarah/Paper.tex

@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Title}
+\icmltitlerunning{Using Bayesian Networks to Predict Change Sets}
 
@@
 \twocolumn[
-\icmltitle{Using Bayesian networks to predict changes}
+\icmltitle{Using Bayesian Networks to Predict Change Sets}
 
-\icmlauthor{Sarah Nadi}{snadi@uwaterloo.ca}
-\icmladdress{University of Waterloo,
+\centering
+\icmlauthor{Sarah Nadi\\}{snadi@uwaterloo.ca\\}
+\icmladdress{University of Waterloo,\\
             Waterloo, ON, Canada}
@@
 
-Changes
+Change is inevitable in any Information Technology(IT) system. New features are added, different configurations are used, upgrades are introduced and new
+software and hardware are added. In a large system, such changes can cause other parts of the system to malfunction without the ability of the analyst to
+foresee this side effect. Accordingly, we need to ensure that we consider the impact of changes before they are implemented, and make sure we adjust the
+potentially affected components accordingly while implementing the change. This leads to the process of change set detection which is finding all the
+components of the system that need to be included in your change set. The change set is the set of components that need to be changed to avoid any side
+effects.
+
+Many of the organizations with large and sophisticated IT systems employ a Configuration Management Database (CMDB) to help them manage these systems.
+The CMDB keeps track of all the components in a system, how they are related, as well as all the changes that have occurred to them. Any component in the CMDB
+is called a Configuration Item (CI). In previous work~\cite{nadi2010}, we mined the CMDB repository for historical co-changes, and used the mined correlations
+to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well. We obtained really
+promising results in terms of recall and precision (69.8\% and 88.5\% respectively).
+
+In this paper, we wish to examine the same problem, but from a Bayesian perspective. Instead of looking at pairwise historical co-changes of CI, we would like
+to consider all the changed CIs in the past to deduce the change relations between them. That is, given a set of observations about how different CIs change
+together, can we deduce the relationships between them that would allow us to predict future change sets. Accordingly, we explore the different ways a Bayesian
+network can be constructed from the data we have, and then test the predictions produced by querying this network to examine the obtained recall and precision.
+We hope that we can obtain better results using Bayesian networks since they are not limited to pairwise comparisons.
+
+The rest of this paper is organized as follows. BLA BLA BLA
 
@@
 \subsection{CMDBs and Change Sets}
-A Configuration Management Database (CMDB) is useful in Enterprise IT Management (EITM) since it provides information about the various critical components in a
-system including hardware, software, and services provided by the company. It records the configuration of these items, their change history, their incident
-history, as well as the relationships between them. Each item stored in the CMDB is referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows
-an example CMDB to illustrate the concepts of CIs and relationships. In this example, there are two software applications being hosted by two different web
-servers. Both applications communicate with the same databases through a load balancer. Such a visualization allows an IT analyst to better understand the
-system at hand.
+The repository of information we are using is the Configuration Management Database (CMDB)~\cite{itil}. The CMDB is useful in Enterprise IT Management (EITM)
+since it provides information about the various critical components in a system including hardware, software, and services provided by the company. It records
+the configuration of these items, their change history, their incident history, as well as the relationships between them. Each item stored in the CMDB is
+referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows an example CMDB to illustrate the concepts of CIs and relationships.
 
 A CMDB provides a basis for decision making processes such as Incident Management, Change Management, etc.  In this paper, we focus on the process of Change
-Management, and in particular, on the problem of change set detection .A \textit{change} is the addition, modification, or removal of anything that could affect
-on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change one CI is the system, other CIs that might
+Management, and in particular, on the problem of change set detection. A \textit{change} is the addition, modification, or removal of anything that could affect
+on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change a CI is the system, other CIs that might
 need to be changed as well must be correctly identified. The set of CIs that will need to be modified for the change to be complete is called a change set.
@@
 
-\subsubsection{Tools}
+BG info and what a bayesian network will look like in our case.
 
-\subsubsection*{Weka}
+\subsection{Bayesian Network Tools}
 
-\subsubsection*{Banjo}
-Banjo is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms: Greedy and
-Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
+\subsubsection{Weka}
+
+WEKA~\cite{hall2009weka} is a toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
+learning problems. It also provides Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
+Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. We will mainly be using WEKA for learning the CPTs in a fixed
+structure network using the SimpleEstimator algorithm~\cite{witten2005data} implemented there. This simply estimates the probability values on edges based on
+the frequency values of each variable given its parents in the training set.
+
+\subsubsection{Banjo}
+Banjo~\cite{banjotool} is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms:
+Greedy and Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
 which basically proposes a random addition or removal of an edge. The other is ``proposeAllLocalMoves'' which proposes all possible moves, and only keeps the
@@
 
-\subsubsection*{JavaBayes}
+\subsubsection{JavaBayes}
+
+JavaBayes is a Java tool for calculating marginal probabilities and expectations in a Bayesian network. In a network, nodes can be set to be observed, and then
+the posterior probability of the remaining nodes can be calculated using Bayes theorem. This is the functionality we need here since the CI that will be
+changed will be considered as observed to be ``true'', and then we need to calculate the posterior of the remaining CIs given this observation.
+
+\section{Constructing the Bayesian Network}
+\label{sec:modelsused}
+
+Before constructing the Bayesian Network, we first had to process and prepare the data we have. Then, in order to construct a Bayesian network to use for
+predictions there are two steps involved. First, determining the structure of the actual network, and then estimating the Conditional Probability Tables
+(CPT). Section~\ref{sec:data} first explains the data set available to us, and the data preprocessing involved before building the network.
+Section~\ref{sec:structure} then explains the different techniques we experimented with to build the network structure. Section~\ref{sec:cpt} shows the last
+step to build the network which is estimating the CPTs. Figure~\ref{fig:process} shows this overall process.
+
+\subsection{Data Preprocessing}
+\label{sec:data}
+
+\subsubsection*{Data Size}
+The original data set available to us in the sample CMDB we have is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data
+was infeasible to work with as no tool could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to
+experiment with Bayesian techniques for change set prediction, it is sufficient to choose a small representative subset of the data. Accordingly, in order to be
+able to test things properly, we used observations from three months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a
+short period, there were already 2,841 distinct CIs appearing and 2,229 observations (i.e. change orders). Therefore, we needed to perform further data
+preprocessing.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=8cm]{graphics/constructingmodel.png}
+\caption{Building the Bayesian Network}
+\label{fig:process}
+\end{figure}
+
+First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with less than 12 different change orders in our
+training observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct
+CIs. Then,
+in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB perspective.
+However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is location for'', and ``backs up''. Adding the related
+parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs. Accordingly, we just kept CIs that
+appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use throughout our models for fair
+comparison. Additionally, filtering out CIs meant filtering out some of the change orders that did not have any CIs satisfying our criteria. This led to us
+having 1,305 change orders instead of 2,229 which was a more manageable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305
+observations (change sets).
+
+\subsubsection*{Data Format}
+\label{sec:format}
+
+In the CMDB, a Change Order has several fields including the requester, the assignee, the start date of the change, the description and the change set field
+(called the `Configuration items' fields). Of the fields in a change order, we use only the change set field. The advantage of only using this one field is that
+change sets are easy to extract and easy to understand. In terms of an observation, CIs were either marked as ``true'' or ``false'' indicating whether they
+have changed or not. Each observation initially consisted of all the CIs in our data set marked as ``false''. For each change order in the training set, we
+would
+mark each CI appearing in the ``Configuration Items'' field as true, while all those not appearing were left as ``false''. Therefore, the set of training
+data consisted of an observation entry for each change order with the CIs marked as ``true'' or ``false'' accordingly.
+
+\subsection{Network Structure}
+\label{sec:structure}
+There were different ways in which we could estimate the structure of the network, and so we built a Bayesian Network using each technique.
+
+\subsubsection{Method 1 (Banjo)}
+\label{sec:model1}
+
+For the first method, we used Banjo to learn the structure of the Bayesian network from the training data. We use the Greedy searching algorithm implemented
+there. Banjo produced five different top-scoring networks (having the same score), and we simply chose one of them as the network to be used.
+
+\subsubsection{Method 2 (WEKA Classification)}
+
+Although our research problem as currently defined is not really a classification problem, we want to examine how the network built by the BayesNet classifier
+in Weka will perform. Accordingly, we estimated the structure of the network using the K2 structure learning algorithm implemented in WEKA which is a greedy
+search based learning algorithm. The default maximum number of parents allowed for any node is one (a Naive Bayesian Network). However, we used a maximum of
+five to allow the nodes to be related since there is really no class label in our data set.
+
+\subsubsection{Method 3 (CMDB Relationships)}
+
+For the third method, we used the relations existing in the CMDB to infer the structure of the network. We used two
+variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
+edge in the Bayesian network from A to B (we will call this model 3A). The second was to reverse the direction of the relationship edges in the CMDB. That is,
+if there is an edge from A to B, we will place an edge in the Bayesian network from B to A (we will call this model 3B). There was one problem, however, with
+the generated networks. Two of the CIs had more than 15 parents to them which means that their CPTs will be intractable to compute. For those two CIs, we simply
+removed all their parents to make the computation tractable.
+
+\subsubsection{Method 4 (Missing Data)}
+
+This method is the same as the third one, except that we experimented with the missing data feature in Weka. Weka allows the user to specify missing values
+for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
+'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. Therefore, this model also has
+two parts: model 4A and model 4B where the first uses the CMDB relationships in their same direction, and the second uses the reverse direction. We could not
+do the same thing with Method 1 (which uses Banjo to learn the structure) since Banjo does not support missing data. WEKA MISSING
+
+\subsection{Estimating the CPTs}
+\label{sec:cpt}
+
+After the structure was determined, we used the SimpleEstimator algorithm~\cite{witten2005data} built in Weka to calculate the CPTs for all networks. The
+SimpleEstimator uses equation BLA GET EQUATION After setting the data set to be the observations in our training set, Weka learned the CPTs for each of the
+networks above. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference as will be explained in the next section.
+
+\section{Experiment Setup}
+\label{sec:setup}
+
+This section explains how we tested the predictions of each of the networks in a way that simulates how an analyst would use the tool in reality. We first
+explain how we simulate this behavior, and then explain how we evaluate the predicted CIs. Algorithm~\ref{alg:simulation} summarizes this setup.
+
+\begin{algorithm}[tb]
+   \caption{Generating Predictions using the Bayesian Network}
+   \label{alg:simulation}
+\begin{algorithmic}
+   \STATE {\bfseries Input:} $change order$
+    \STATE {\bfseries Input:} $Bayesian network$
+    \STATE {\bfseries Input:} $threshold$
+    \STATE Set $initial CI$ = first CI in changeOrder
+    \STATE Set $occurred set$ = $change order$ - $initial CI$
+    \STATE Initialize $observed set$ = {$initial CI$}
+    \STATE Initialize $predicted set$ = {}
+   \REPEAT  
+    \STATE Initialize $new predictions$ = {}
+    \FOR{$node$ in $BayesianNetwork$}
+    \STATE $posterior$ = query $node$ in $Bayesian network$ using Banjo given observed set
+    \IF{$posterior > threshold$ and $node$ not in $predicted set$} 
+      \STATE Add $node$ to $new predictions$   
+    \ENDIF
+    \ENDFOR
+
+  \FOR{$prediction$ in $new predictions$}
+    \IF{$prediction$ in  $occurred set$}
+      \STATE Add $prediction$ to $observedSet$
+    \ENDIF
+    \ENDFOR
+\STATE Add $new predictions$ to $predicted set$
+
+   \UNTIL{$new predictions$ is empty}
+\STATE $recall = (predicted set \cap occurred set)/occurred set$
+\STATE $precision = (predicted set \cap occurred set)/predicted set$
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Change Set Detection Process Simulation}
+
+The change set detection process provided is an iterative, collaborative process between the tool and the analyst. When the tool suggests CIs, the analyst can
+accept or reject these CIs (i.e add them to the change set or not). Based on the CIs added to the change set, the analyst can ask the tool for more
+suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''.
+
+To simulate this process for our experiments, we do the following. At this point, we have the Bayesian network ready, and we would like to perform Inference.
+More formally, given that a CI will change (our observation), we want to infer the probability that the other CIs in the network might change as well.
+Unfortunately, neither of the two tools previously used provide a general inference engine. We, therefore, use JavaBayes in this step since it accepts
+the same BIFF format used by Weka. We use a one month testing set where we try to predict all the change orders in April 2008. There was a total of
+883 change orders in that month which we use in our testing set.
+
+For each change order, we would take the first CI as the initial CI to change, then we would set that as an observed node, and update
+the beliefs of all the nodes (CIs) in the network using JavaBayes. We are basically checking the probability of each CI being ``true''. We used different
+cutoff thresholds for the posterior probabilities we would consider. We would then loop on all the updated CIs, and add those that match our threshold criteria
+to
+the predicted change set. For example a threshold of 0.2 means that only nodes that have a posterior probability greater than 0.2 of having the value ``true''
+will be returned. To simulate a real life scenario, we then checked which of these predicted CIs actually lies in the target change set we are trying to
+predict. This is similar to an analyst accepting CIs into their change set. These common CIs would then also be marked as observed so that we can predict
+what else will need to change given that these CIs are also changing. Again, we would calculate the posterior probability, and continue doing
+so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst or not) are part of the predicted
+set.
 
@@
 
-Similar to Hassan et. al~\cite{hassan2004predicting}, we define the \textit{Predicted Set} (P) as the set of all CIs DRACA suggests through the full the
-iteration process (see Figure~\ref{fig:process}). We define the \textit{Occurred Set} (O) as the CIs remaining in the change set after excluding the Initial CI
+Similar to Hassan et. al~\cite{hassan2004predicting}, wescore define the \textit{Predicted Set} (P) as the set of all predicted. We define the \textit{Occurred
+Set} (O) as the CIs remaining in the change set after excluding the Initial CI
 provided by the analyst (i.e Change Set - Initial CI). The intersection of the predicted set and the occurred set, called $PO$, is the common CIs in both sets.
@@
 
+\section{Results}
+\label{sec:results}
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=8cm]{graphics/cmdbresults.png}
+\caption{F-measure of the different Bayesian networks produced using the CMDB relationships}
+\label{fig:cmdbresults}
+\end{figure}
+
+For all the networks generated from the four different methods, and following the procedure described in Section~\ref{sec:setup}, we calculated the average
+recall and precision for all the change orders in the test set. Based on the average recall and precision values, we calculated the F-measure. For better
+visualization, we plot the results from the first two methods together, and the results from the second two methods together.
+
+Figure~\ref{fig:cmdbresults} shows the F-measure obtained from the four different networks built based on CMDB relationships for structural information. The
+first observation is that reversing the relationship edges did not produce any difference in the results. The curves for the CMDB relationships and the reversed
+CMDB relationships are completely overlapping. The same things applies to the curves of CMDB relationships with missing data, and reversed CMDB relationships
+with missing data. The second observation here is that denoting CIs that did not appear in the change order as missing data rather than giving them the value
+``false'' produced slightly better results. However, the main conclusion from these results is that given each change order, all of the predicted CIs in the
+network already had a posterior probability of greater than 0.9. That is why the F-measure is the same for all thresholds (excluding 0), and only
+changes when the threshold becomes 0.9. This seems rather strange, and despite the relatively high F-measure, it does not seem that the structure
+of these networks is reflective of their relationships in terms of change.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=8cm]{graphics/banjowekares.png}
+\caption{F-measure from network structures estimated by WEKA versus Banjo}
+\label{fig:wekabnj}
+\end{figure}
+
+Figure~\ref{fig:wekabnj} compares the F-measure
+
 \section{Related Work}
@@
 element, they also add features such as authors, change significance levels etc. and try to predict if two elements are co-changes or not accordingly. Thus,
-their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as co-changes or not.
+their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as a co-change or not.
 They use the K2 algorithm proposed by Cooper et. al~\cite{cooper1992bayesian} to estimate the structure of the Bayesian network, and use the SimpleEstimator
 algorithm built in WEKA~\cite{witten2005data}.
-
-\section{Models Used}
-\label{sec:modelsused}
-
-In order to construct a Bayesian network which we can use to make predictions, there were two steps involved. First, determining the structure of the actual
-network, and then estimating the Conditional Probability Tables (CPT). There were different ways in which we could estimate the structure of the network, and
-so we built a model for each technique. However, for all models, we used the SimpleEstimator algorithm~\cite{written2005data} built in Weka to calculate the
-CPTs.
-
-Before building any of the models, some data preprocessing was necessary. We first describe this preprocessing, and then proceed to explain each of the FOUR
-models built.
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=7cm]{graphics/constructingmodel.png}
-\caption{Generating the Bayesian Network}
-\label{fig:process}
-\end{figure}
-
-\subsubsection{Data Preprocessing}
-The original data set has 7,999 distinct CIs, and 27,305 change orders. This amount of data was infeasible to work with as no tool could handle such a large
-amount of variables in a Bayesian network. For the purposes of this project, which is mainly to experiment with Bayesian techniques for change set prediction,
-it is sufficient to choose a small representative subset of the data. Accordingly, in order to be able to test things properly, we used observations from three
-months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a short period, there were already 2,841 distinct CIs appearing and
-2,229 observations (i.e. change orders). Therefore, we needed to perform further data preprocessing.
-
-First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with 12 different change orders in our training
-observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct CIs. Then,
-in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB perspective.
-However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is location for'', and ``backs up''. Adding the related
-parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs. Accordingly, we just kept CIs tht
-appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use throughout our models for fair
-comparison. Additionally, filtering out CIs meant filtering out some of the observations that did not have any CIs satisfying our criteria. This led to us
-having 1,305 observations instead of 2,229 which was a more manegeable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305
-observations (change sets).
-
-DESCRIBE GENERATED OBSERVATIONS FILE
-
-FIGURE FOR DATA PROCESSING
-
-\subsection{Model 1 (Banjo)}
-\label{sec:model1}
-
-For the first model, we used Banjo to learn the structure of the Bayesian network from the training data which used a Greedy searching algorithm. Banjo
-produced five different top-scoring networks (having the same score), and we simply chose one of them as the network to be used. This network was then
-formatted as an BIF XML file, and was fed into Weka's network analyzer tool. After setting the data set to be the observations in our training set, Weka
-learned the CPTs for the network produced by Banjo. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference.
-
-\subsection{Model 2 (CMDB Relationships)}
-
-For the second model, we used the relations existing in the CMDB to infer the structure of the network (that is place the edges between the nodes). We used two
-variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
-edge in the Bayesian network from A to B (we will call this model 2A). The second was to reverse the direction of the relationship edges in the CMDB. That is,
-if there is an edge from A to B, we will place an edge in the Bayesian network from B to A (we will call this model 2B). In both cases, however, we chose to
-ignore some relationship edges (WHY, TRY WITHOUT). There was one problem, however, with the generated network. Two of the CIs had more than 15 parents to them
-which means that their CPTs will be intractable to compute. For those two CIs, we simply removed all their parents to make the computation tractable.
-
-\subsection{Model 3 (Missing Data)}
-
-This model is the same as the second model, except that we experimented with the missing data feature in Weka. Weka allows the user to specify missing values
-for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
-'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. Therefore, this model also has
-two parts: model 3A and model 3B where the first uses the CMDB relationships in their same direction, and the second uses the reverse direction.
-
-\begin{algorithm}[tb]
-   \caption{Bubble Sort}
-   \label{alg:example}
-\begin{algorithmic}
-   \STATE {\bfseries Input:} changeOrder
-    \STATE {\bfseries Input:} BayesianNetwork
-    \STATE {\bfseries Input:} threshold
-    \STATE Set initialCI = first CI in changeOrder
-    \STATE Set occurredSet = changeOrder - initialCI
-    \STATE Initialize observedSet = {firstCI}
-    \STATE Initialize predictedSet = {}
-   \REPEAT  
-    \STATE Initialize newPredictions = {}
-    \FOR{$node$ {\bfseries to} $BayesianNetwork$}
-    \STATE posterior = perform inference using Banjo
-    \IF{$posterior > threshold$} 
-      \STATE Add node to newPredictions   
-    \ENDIF
-    \ENDFOR
-
-  \FOR{$prediction$ {\bfseries to} $newPredictions$}
-    \IF{$prediction  occurredSet$}
-      \STATE Add prediction to observedSet
-    \ENDIF
-    \ENDFOR
-\STATE Add newPredictions to predictedSet
-
-   \UNTIL{$newPredictions$ is $empty$}
-\end{algorithmic}
-\end{algorithm}
-
-\section{Experimental Work}
-\label{sec:exp}
-
-\subsubsection*{Inference}
-
-At this point, we have the Bayesian network ready, and we would like to perform Inference. More formally, given that a CI will change (our evidence), we want
-to infer the probability that the other CIs in the network might change as well. Unfortunately, neither of the two tools previously used provide an Bayesian
-inference engine. We, therefore, use JavaBayes in this step since it accepts the same ARFF format used by Weka. We using a one month testing set where we try
-to predict all the change orders in April 2008 (883 reports). There was a total of XX change orders in that month. For each change order, we would take the
-first CI as the initial CI to change, then we would set that as evidence, and update the beliefs of all the nodes (CIs) in the network using JavaBayes. We would
-then loop on all the updated CIs, and add those that match our threshold criteria to the predicted change set. To simulate a real life scenario, we then checked
-which of these predicted CIs actually lies in the target change set we are trying to predict. This is similar to an analyst accepting CIs into their change set.
-These common CIs would then also be marked as observations so that we can predict now that I'm going to change B and C too, what else do I need to change.
-Again, we would calculate the posterior probability, and continue doing so until there are no more common CIs. We would then calculate the recall, precision,
-and F-measure accordingly.
 

### 2010-12-08 trunk/cs886-sarah/Paper.tex

@@
 
-BG info and what a bayesian network will look like in our case.
+Bayesian Networks, first introduced by Pearl~\cite{pearl1988probabilistic} are probabilistic graphical models. Each node in a graph represents a random
+variable, and nodes are connected through edges which represent probabilistic dependencies~\cite{benbayesian}. A Bayesian network is a Directed
+Acyclic Graph (DAG), and is essentially defined by two aspects: the structure of the graph and the Conditional Probability Tables (CPTs) probabilistically
+relating connected nodes. In this work, a node in the Bayesian network would be a CI which can take values {``true'',``false''} to donate the probability of it
+changing or not. The CPTs would essentially give the probability of one CI changing given its parents.
+
+Given a set of data consisting of observations for our variable space, machine learning techniques can be used to estimate the network. If the structure of the
+network is known, and we have full observability (i.e we can observe values for each variable), then learning the CPTs can be achieved through
+Maximum-likelihood estimation (MLE)~\cite{benbayesian}. MLE tries to maximize the log likelihood of the the training data which can simply be done by
+calculating the frequency of occurrence of the values of each variable given its parents. If the structure is known, but we only have partial observability (i.e
+missing data), then we can use a variation of MLE which is the Expectation Maximization algorithm~\cite{dempster1977maximum}. If the structure is not known in
+the first place, then searching through the model space, and calculating network scores for each produced network is necessary. The simplest algorithm for this
+is a Greedy search algorithm~\cite{friedman1996learning}.
+
+While searching for the best network that fits the training data, scores are needed for the produced networks in order to know which one is the best one. There
+are several types of scores that have been proposed in the literature. Two of the commonly used scores are the Minimum Description Length (MDL)
+score~\cite{rissanen1986stochastic} and the BDe score~\cite{heckerman1995learning}. The MDL score is based on finding the length of data required to store a
+network. It tries to balance the complexity of the induced network with the degree of accuracy of which the network represents the frequencies in the training
+data. Therefore, the lower the MDL score of a network, the better it is. The BDe score is proportional to the posterior probability of each network structure
+being scored, given the data. Accordingly, we are searching for the network that maximizes this probability. For large data sets, both metrics score
+structures similarly~\cite{friedman1996learning}. Since they are similar, we simply chose to use the BDe score in this work to compare the produced network
+structures.
 
@@
 Banjo~\cite{banjotool} is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms:
-Greedy and Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
-which basically proposes a random addition or removal of an edge. The other is ``proposeAllLocalMoves'' which proposes all possible moves, and only keeps the
-best one (?). Banjo uses the BDe metric to compute a network's score.
+Greedy and Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. Banjo uses the BDe metric to compute a
+network's score. We used the Greedy search algorithm in our experiments.
 
@@
 variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
-edge in the Bayesian network from A to B (we will call this model 3A). The second was to reverse the direction of the relationship edges in the CMDB. That is,
-if there is an edge from A to B, we will place an edge in the Bayesian network from B to A (we will call this model 3B). There was one problem, however, with
+edge in the Bayesian network from A to B. The second was to reverse the direction of the relationship edges in the CMDB. That is,
+if there is an edge from A to B, we will place an edge in the Bayesian network from B to A. There was one problem, however, with
 the generated networks. Two of the CIs had more than 15 parents to them which means that their CPTs will be intractable to compute. For those two CIs, we simply
 removed all their parents to make the computation tractable.
-
-\subsubsection{Method 4 (Missing Data)}
-
-This method is the same as the third one, except that we experimented with the missing data feature in Weka. Weka allows the user to specify missing values
-for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
-'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. Therefore, this model also has
-two parts: model 4A and model 4B where the first uses the CMDB relationships in their same direction, and the second uses the reverse direction. We could not
-do the same thing with Method 1 (which uses Banjo to learn the structure) since Banjo does not support missing data. WEKA MISSING
 
@@
 networks above. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference as will be explained in the next section.
+
+Additionally, we experimented with the missing data feature in WEKA. Weka allows the user to specify missing values
+for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
+'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. We tried learning the CPTs for
+the network produced by WEKA classification, and for that produced using the CMDB relationships. The missing values were only used to learn the CPTs except in
+the case for WEKA where the structure was also learned using the missing values.
 
@@
 
-For all the networks generated from the four different methods, and following the procedure described in Section~\ref{sec:setup}, we calculated the average
+For all the networks generated from different methods explained in Section~\ref{sec:structure}, and following the procedure described in
+Section~\ref{sec:setup}, we calculated the average
 recall and precision for all the change orders in the test set. Based on the average recall and precision values, we calculated the F-measure. For better
-visualization, we plot the results from the first two methods together, and the results from the second two methods together.
+visualization, we plot the results from all the networks based on CMDB structure together, and the results from those based on Banjo and WEKA together.
 
@@
 
-Figure~\ref{fig:wekabnj} compares the F-measure
+Figure~\ref{fig:wekabnj} compares the F-measure obtained from the graphs whose structure was built using Banjo versus those built using WEKA.
 

### 2010-12-09 trunk/cs886-sarah/Paper.tex
missing final runs for results isA

@@
-\documentclass{article}
+\documentclass[10pt,twocolumn,letterpaper]{article}
 
-% For figures
+\usepackage{cvpr}
+\usepackage{times}
 \usepackage{graphicx}
-
-% For citations
-\usepackage{natbib}
-
-% For algorithms
 \usepackage{algorithm}
@@
 \usepackage{url}
-
-% As of 2010, we use the hyperref package to produce hyperlinks in the
-% resulting PDF.  If this breaks your system, please commend out the
-% following usepackage line and replace \usepackage{icml2010} with
-% \usepackage[nohyperref]{icml2010} above.
 \usepackage{hyperref}
 
-% Packages hyperref and algorithmic misbehave sometimes.  We can fix
-% this with the following command.
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-\usepackage[accepted]{icml2010}
+\cvprfinalcopy % *** Uncomment this line for the final submission
 
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Using Bayesian Networks to Predict Change Sets}
+\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
+\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
+
+% Pages are numbered in submission mode, and unnumbered in camera-ready
+\ifcvprfinal\pagestyle{empty}\fi
 
@@
 
-\twocolumn[
-\icmltitle{Using Bayesian Networks to Predict Change Sets}
+\title{Using Bayesian Networks to Predict Change Sets in an IT System}
 
-\centering
-\icmlauthor{Sarah Nadi\\}{snadi@uwaterloo.ca\\}
-\icmladdress{University of Waterloo,\\
-            Waterloo, ON, Canada}
-
-\vskip 0.3in
-]
+\author{Sarah Nadi\\snadi@uwaterloo.ca\\University of Waterloo\\ Ontario, Canada}        
+\thispagestyle{empty}
+\maketitle
 
 \begin{abstract} 
-STILLLL
+When change one component in an IT system, an analyst must be careful not to negatively impact any of the other system components. Detecting the complete
+change set (all those components that need to be changed to avoid any side effects) is therefore very important. Given historic information about how different
+components of a system change together, we build a Bayesian network that can be used to predict this change set. Given an initial component to be changed in
+the system, the posterior probability of the other components changing can be calculated using the network. In this paper, we experiment with different
+techniques for building the Bayesian network based on the data available, and report on their predictive power in terms of recall and precision.
 \end{abstract}
@@
 In this paper, we wish to examine the same problem, but from a Bayesian perspective. Instead of looking at pairwise historical co-changes of CI, we would like
-to consider all the changed CIs in the past to deduce the change relations between them. That is, given a set of observations about how different CIs change
+to consider all the changed CIs, and learn the relationships between them. That is, given a set of observations about how different CIs change
 together, can we deduce the relationships between them that would allow us to predict future change sets. Accordingly, we explore the different ways a Bayesian
@@
 
-The rest of this paper is organized as follows. BLA BLA BLA
+The rest of this paper is organized as follows. Section~\ref{sec:bg} provides the background information necessary for this work. It explains the nature of
+CMDBs, Bayesian networks, and the tools we used for analyzing Bayesian networks. Section~\ref{sec:constructing} then presents the techniques we used to build
+our Bayesian network. This includes how we process the data, and build the different Bayesian networks accordingly. Section~\ref{sec:setup} details the
+setup we use for our experiments, and Section~\ref{sec:results} provides the results obtained. Section~\ref{sec:discussion} provides a general insight to the
+results, and future work that could be done to improve them. Section~\ref{sec:concl} concludes this paper.
 
 \section{Background}
+\label{sec:bg}
 
@@
 on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change a CI is the system, other CIs that might
-need to be changed as well must be correctly identified. The set of CIs that will need to be modified for the change to be complete is called a change set.
+need to be changed as well must be correctly identified. We refer to the set of CIs that will need to be modified for the change to be complete ias the
+\textit{change set}.
 
@@
 
-Bayesian Networks, first introduced by Pearl~\cite{pearl1988probabilistic} are probabilistic graphical models. Each node in a graph represents a random
+Bayesian Networks, first introduced by Pearl~\cite{pearl1988probabilistic} are a type of probabilistic graphical models. Each node in the network
+represents a random
 variable, and nodes are connected through edges which represent probabilistic dependencies~\cite{benbayesian}. A Bayesian network is a Directed
 Acyclic Graph (DAG), and is essentially defined by two aspects: the structure of the graph and the Conditional Probability Tables (CPTs) probabilistically
-relating connected nodes. In this work, a node in the Bayesian network would be a CI which can take values {``true'',``false''} to donate the probability of it
-changing or not. The CPTs would essentially give the probability of one CI changing given its parents.
+relating connected nodes. In our work, a node in the Bayesian network would be a CI which can take values {``true'',``false''} to donate the probability of it
+changing or not. The CPTs would essentially give the probability of one CI changing given its parents' values.
 
@@
 Maximum-likelihood estimation (MLE)~\cite{benbayesian}. MLE tries to maximize the log likelihood of the the training data which can simply be done by
-calculating the frequency of occurrence of the values of each variable given its parents. If the structure is known, but we only have partial observability (i.e
-missing data), then we can use a variation of MLE which is the Expectation Maximization algorithm~\cite{dempster1977maximum}. If the structure is not known in
+calculating the frequency of occurrence of the values of each variable given its parents. If the structure is known, but we only have partial observability
+(i.e., missing data), then we can use a variation of MLE which is the Expectation Maximization algorithm~\cite{dempster1977maximum}. If the structure is not
+known in
 the first place, then searching through the model space, and calculating network scores for each produced network is necessary. The simplest algorithm for this
@@
 being scored, given the data. Accordingly, we are searching for the network that maximizes this probability. For large data sets, both metrics score
-structures similarly~\cite{friedman1996learning}. Since they are similar, we simply chose to use the BDe score in this work to compare the produced network
-structures.
+structures similarly~\cite{friedman1996learning}.
 
@@
 
-\subsubsection{Weka}
+There are many machine learning and Bayesian analysis tools available. In this work, we use three of these tools which we briefly describe
+below.
 
-WEKA~\cite{hall2009weka} is a toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
-learning problems. It also provides Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
-Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. We will mainly be using WEKA for learning the CPTs in a fixed
-structure network using the SimpleEstimator algorithm~\cite{witten2005data} implemented there. This simply estimates the probability values on edges based on
+\subsubsection{WEKA}
+
+WEKA~\cite{hall2009weka} is a Java toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
+learning problems. It has Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
+Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. IN terms of learning CPTs, one of the algorithms WEKA implements is
+the SimpleEstimator algorithm~\cite{witten2005data} implemented which we use in our work. This simply estimates the probability values on edges based on
 the frequency values of each variable given its parents in the training set.
@@
 Banjo~\cite{banjotool} is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms:
-Greedy and Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. Banjo uses the BDe metric to compute a
-network's score. We used the Greedy search algorithm in our experiments.
+Greedy and Simulated Annealing. Banjo uses the BDe metric to compute a network's score, and produces the highest scoring network found. We used the Greedy
+search algorithm in our experiments.
+
+PIC STILL NEEDS WORK
+
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=15cm]{graphics/constructingmodel.png}
+\caption{Building the Bayesian Network}
+\label{fig:process}
+\end{figure*}
 
@@
 \section{Constructing the Bayesian Network}
-\label{sec:modelsused}
+\label{sec:constructing}
 
-Before constructing the Bayesian Network, we first had to process and prepare the data we have. Then, in order to construct a Bayesian network to use for
+Before constructing the Bayesian Network, we first have to process and prepare the data we have. In order to construct a Bayesian network to use for
 predictions there are two steps involved. First, determining the structure of the actual network, and then estimating the Conditional Probability Tables
@@
 \subsubsection*{Data Size}
-The original data set available to us in the sample CMDB we have is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data
+The data set available to us in the sample CMDB we have is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data
 was infeasible to work with as no tool could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to
@@
 
-\begin{figure}[!t]
-\centering
-\includegraphics[width=8cm]{graphics/constructingmodel.png}
-\caption{Building the Bayesian Network}
-\label{fig:process}
-\end{figure}
-
 First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with less than 12 different change orders in our
 training observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct
-CIs. Then,
-in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB perspective.
-However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is location for'', and ``backs up''. Adding the related
-parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs. Accordingly, we just kept CIs that
-appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use throughout our models for fair
-comparison. Additionally, filtering out CIs meant filtering out some of the change orders that did not have any CIs satisfying our criteria. This led to us
-having 1,305 change orders instead of 2,229 which was a more manageable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305
-observations (change sets).
+CIs. Then, in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB
+perspective. Adding the related parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs.
+Accordingly, we just kept CIs that appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use
+throughout our models for fair comparison. Additionally, filtering out CIs meant filtering out some of the change orders that did not have any CIs satisfying
+our criteria (i.e. all the CIs that changed in them were outside our 170 CI set). This led to us having 1,305 change orders instead of 2,229 which was more
+manageable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305 observations (change sets).  TRAINING?
 
@@
 \label{sec:structure}
-There were different ways in which we could estimate the structure of the network, and so we built a Bayesian Network using each technique.
+There were essentially three different ways in which we could estimate the structure of the network. We, thus, built a Bayesian Network using each technique to
+be able to compare their performance.
 
@@
 Although our research problem as currently defined is not really a classification problem, we want to examine how the network built by the BayesNet classifier
-in Weka will perform. Accordingly, we estimated the structure of the network using the K2 structure learning algorithm implemented in WEKA which is a greedy
+in WEKA will perform. Accordingly, we estimated the structure of the network using the K2 structure learning algorithm implemented in WEKA which is also a
+greedy
 search based learning algorithm. The default maximum number of parents allowed for any node is one (a Naive Bayesian Network). However, we used a maximum of
-five to allow the nodes to be related since there is really no class label in our data set.
+five to allow the nodes to be related since there is really no class label in our data set. We could not experiment with other techniques in WEKA because they
+do not allow multiple parents.
 
@@
 For the third method, we used the relations existing in the CMDB to infer the structure of the network. We used two
-variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
+variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is, if there is a relationship edge in the CMDB from A to
+B, we will place an
 edge in the Bayesian network from A to B. The second was to reverse the direction of the relationship edges in the CMDB. That is,
-if there is an edge from A to B, we will place an edge in the Bayesian network from B to A. There was one problem, however, with
+if there is a relationship edge from A to B in the CMDB, we will place an edge in the Bayesian network from B to A. There was one problem, however, with
 the generated networks. Two of the CIs had more than 15 parents to them which means that their CPTs will be intractable to compute. For those two CIs, we simply
@@
 
-After the structure was determined, we used the SimpleEstimator algorithm~\cite{witten2005data} built in Weka to calculate the CPTs for all networks. The
-SimpleEstimator uses equation BLA GET EQUATION After setting the data set to be the observations in our training set, Weka learned the CPTs for each of the
+After the structure was determined, we used the SimpleEstimator algorithm~\cite{witten2005data} built in WEKA to calculate the CPTs for all networks. After
+setting the data set to be the observations in our training set (an ARFF file), WEKA learned the CPTs for each of the
 networks above. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference as will be explained in the next section.
 
-Additionally, we experimented with the missing data feature in WEKA. Weka allows the user to specify missing values
+Additionally, we experimented with the missing data feature in WEKA. WEKA allows the user to specify missing values
 for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
-'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. We tried learning the CPTs for
-the network produced by WEKA classification, and for that produced using the CMDB relationships. The missing values were only used to learn the CPTs except in
-the case for WEKA where the structure was also learned using the missing values.
+'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. When using Banjo or the CMDB
+relationships to estimate the structure of the network, the missing values were only inserted when learning the CPTs. When using WEKA, the missing values were
+inserted both to learn the network, and to estimate the CPTs.
 
@@
 This section explains how we tested the predictions of each of the networks in a way that simulates how an analyst would use the tool in reality. We first
-explain how we simulate this behavior, and then explain how we evaluate the predicted CIs. Algorithm~\ref{alg:simulation} summarizes this setup.
+explain how we simulate this behavior, and then explain how we evaluate the predicted CIs. Algorithm~\ref{alg:simulation} summarizes these steps taken to
+predict a change order.
 
@@
 accept or reject these CIs (i.e add them to the change set or not). Based on the CIs added to the change set, the analyst can ask the tool for more
-suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''.
+suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''. This process is explained in more
+details by Nadi et al.~\cite{nadi2010}.
 
 To simulate this process for our experiments, we do the following. At this point, we have the Bayesian network ready, and we would like to perform Inference.
-More formally, given that a CI will change (our observation), we want to infer the probability that the other CIs in the network might change as well.
+More formally, given that a CI will change (our observation), we want to update our beliefs that the other CIs in the network might change as well.
 Unfortunately, neither of the two tools previously used provide a general inference engine. We, therefore, use JavaBayes in this step since it accepts
-the same BIFF format used by Weka. We use a one month testing set where we try to predict all the change orders in April 2008. There was a total of
-883 change orders in that month which we use in our testing set.
+the same XML BIFF format produced by WEKA. [We use a one month testing set where we try to predict all the change orders in April 2008. There was a total of
+883 change orders in that month which we use in our testing set. ] - REMOVE?
 
-For each change order, we would take the first CI as the initial CI to change, then we would set that as an observed node, and update
-the beliefs of all the nodes (CIs) in the network using JavaBayes. We are basically checking the probability of each CI being ``true''. We used different
+For each change order, we would take the first CI recorded as the initial CI to change, then we would set that as an observed node, and update
+the beliefs of all the nodes (CIs) in the network using JavaBayes. We are basically checking the posterior probability of each CI being ``true'' given our
+observation. We used different
 cutoff thresholds for the posterior probabilities we would consider. We would then loop on all the updated CIs, and add those that match our threshold criteria
@@
 predict. This is similar to an analyst accepting CIs into their change set. These common CIs would then also be marked as observed so that we can predict
-what else will need to change given that these CIs are also changing. Again, we would calculate the posterior probability, and continue doing
+what else will need to change given that these CIs are also changing. Again, we would calculate the posterior probability given the new observations, and
+continue doing
 so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst or not) are part of the predicted
-set.
+set. These steps (shown in Algorithm~\ref{alg:simulation} would be repeated for each change order.
 
@@
 
-We need a way to evaluate the predicted CIs. The recall and precision measures from the information retrieval field are appropriate for this type of evaluation.
+We need a way to evaluate the predicted CIs. The recall and precision measures from Information Retrieval are appropriate for this type of evaluation.
 Recall measures the proportion of correct CIs retrieved by the system, while precision measures the proportion of suggested CIs that are correct~\cite{van79}.
 
-Similar to Hassan et. al~\cite{hassan2004predicting}, wescore define the \textit{Predicted Set} (P) as the set of all predicted. We define the \textit{Occurred
+Similar to Hassan et. al~\cite{hassan2004predicting}, we define the \textit{Predicted Set} (P) as the set of all predicted. We define the \textit{Occurred
 Set} (O) as the CIs remaining in the change set after excluding the Initial CI
@@
 
+\begin{table*}[h!t!]
+\centering
+\resizebox{\textwidth }{!}{
+\begin{tabular}{ c | c |c ||c | c || c |c ||c | c || c |c ||c| c||c|c||c|c}
+\textbf{Threshold} 
+& \multicolumn{2}{|c||}{\textbf{Banjo}} 
+& \multicolumn{2}{|c||}{\textbf{Banjo Missing}}
+&\multicolumn{2}{|c||}{\textbf{WEKA}}
+&\multicolumn{2}{|c||}{\textbf{WEKA Missing}}
+&\multicolumn{2}{|c||}{\textbf{CMDB}}
+&\multicolumn{2}{|c||}{\textbf{CMDB Missing}}
+&\multicolumn{2}{|c||}{\textbf{Reversed CMDB}} 
+&\multicolumn{2}{|c}{\textbf{Reversed CMDB Missing}}\\
+\hline
+ & \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall}
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}\\
+\hline\hline
+0&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473\\
+\hline
+0.1&0.649&0.493&0.715&0.473&0.707&0.501&0.715&0.473&0.569&0.470&0.715&0.473&0.583&0.471&0.715&0.473\\
+\hline
+0.2&0.643&0.501&0.715&0.473&0.704&0.509&0.715&0.473&0.568&0.470&0.715&0.473&0.580&0.470&0.715&0.473\\
+\hline
+0.3&0.634&0.510&0.715&0.473&0.700&0.519&0.715&0.473&0.568&0.469&0.715&0.473&0.569&0.469&0.715&0.473\\
+\hline
+0.4&0.624&0.511&0.715&0.473&0.696&0.524&0.715&0.473&0.568&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.5&0.610&0.500&0.715&0.473&0.682&0.528&0.715&0.473&0.568&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.6&0.607&0.500&0.715&0.473&0.674&0.529&0.715&0.473&0.567&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.7&0.595&0.499&0.715&0.473&0.646&0.526&0.715&0.473&0.567&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.8&0.595&0.499&0.715&0.473&0.627&0.526&0.715&0.473&0.567&0.469&0.715&0.473&0.567&0.469&0.714&0.473\\
+\hline
+0.9&0.574&0.475&0.715&0.473&0.588&0.489&0.567&0.469&0.567&0.469&0.715&0.473&0.567&0.469&0.712&0.473\\
+\hline
+1&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000\\
+\end{tabular}
+}
+\centering
+\caption{Summary of Results}
+\label{tab:results}
+\end{table*}
+
 \section{Results}
@@
 
-For all the networks generated from different methods explained in Section~\ref{sec:structure}, and following the procedure described in
-Section~\ref{sec:setup}, we calculated the average
-recall and precision for all the change orders in the test set. Based on the average recall and precision values, we calculated the F-measure. For better
-visualization, we plot the results from all the networks based on CMDB structure together, and the results from those based on Banjo and WEKA together.
+For all the networks generated from the different methods explained in Section~\ref{sec:structure}, and following the procedure described in
+Section~\ref{sec:setup}, we calculated the average recall and precision for all the change orders in the test set. Based on the average recall and precision
+values, we calculated the F-measure. Table~\ref{tab:results} summarizes the recall and precision obtained for all models. For better visualization, we plot the
+results from all the networks based on CMDB structure together, and the results from those based on Banjo and WEKA together.
 
@@
 
-Figure~\ref{fig:wekabnj} compares the F-measure obtained from the graphs whose structure was built using Banjo versus those built using WEKA.
+Figure~\ref{fig:wekabnj} compares the F-measure obtained from the graphs whose structure was built using Banjo versus those built using WEKA. It also shows the
+effects including missing data in both cases. Surprisingly, it seems that the network built using WEKA has better performance (although only slightly better).
+However, marking CIs that have not been changed as unobserved (missing) instead of ``false'' slightly deteriorated the predictive performance. Overall, the
+performance of the two networks are very similar which comes as a surprise since WEKA is really a classification engine and assumes that the last variable in
+the data is the target class which is not the case in our data.
+
+We also note when the threshold is 1.0 (i.e no predictions are being made), the recall is not 0 although the precision is 1. This is because there are many
+change sets of size 1 which makes the occurred set of size 0, and thus recall being automatically 1. Please refer to Nadi et. al~\cite{nadi2010} for more
+details. Additionally, all networks start at the same recall and precision values when the threshold is 0 since simply all 170 CIs are predicted. This is
+because in the Bayesian network, the probability of any CI is never really zero. It may be very close to zero, but will never be completely unlikely.
 
@@
 
-Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. However, their work is on the level of source code changes. They build three
+Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. They build three
 different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
@@
 Maximization (EM) algorithm~\cite{dempster1977maximum}. The second was solely based on historic information where the network is build using a greedy structure
- learning algorithm~\cite{friedman1996learning}. They did some preprocessing to their data such as filtering out large changes (with more than 30 elements
-changed at once) since this was probably an insignificant change.
+ learning algorithm~\cite{friedman1996learning}. The main difference between their work and ours is that their work is on the level of source
+code while ours is on a system level.
 
@@
 
+\section{Future Work}
+\label{sec:future}
+
+There is still some work that needs to be done in this area. First, it would be intersting to experiment with undirected graphs. For our problem, when two CIs,
+A and B, occur in the same change order, it is hard to determine whether A caused B to change or B caused A to change. Therefore, using an undirected model,
+such as a Markov network could be interesting in this setting. Unfortunately, due to time constraints, we could not experiment with that. Another possible
+expansion is to use the CMDB relationships more carefully when determining the structure of the network. Currently, we either add edges to the Bayesian network
+following the direction of those in the CMDB or in the reverse direction. It would be interesting to determine which set of relationships semantically make
+sense to have the causality relationship in their same direction, and which make sense to have it in the opposite direction. That way, the structure of the
+network produced could perhaps be more accurate.
+
 \section{Conclusion}
@@
 
+In this work, we experimented with Bayesian networks in the context of change set predictions for IT systems. Given the CIs the analyst wants to change in the
+system, we can predict what other CIs might need to be changed as well. We built different Bayesian networks based on different information such as CMDB
+relationships or using structural learning algorithms such as the Greedy algorithm.
+
 \bibliography{references}
-\bibliographystyle{icml2010}
+\bibliographystyle{ieee}
 

### 2010-12-10 cs886-sarah/Paper.tex
semi final draft isA

@@
 \usepackage{url}
-
-% As of 2010, we use the hyperref package to produce hyperlinks in the
-% resulting PDF.  If this breaks your system, please commend out the
-% following usepackage line and replace \usepackage{icml2010} with
-% \usepackage[nohyperref]{icml2010} above.
 \usepackage{hyperref}
 
-% Packages hyperref and algorithmic misbehave sometimes.  We can fix
-% this with the following command.
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-\usepackage[accepted]{icml2010}
+\cvprfinalcopy % *** Uncomment this line for the final submission
 
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Title}
+\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
+\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
+
+% Pages are numbered in submission mode, and unnumbered in camera-ready
+\ifcvprfinal\pagestyle{empty}\fi
 
@@
 
-\twocolumn[
-\icmltitle{Using Bayesian networks to predict changes}
+\title{Using Bayesian Networks to Predict Change Sets in an IT System}
 
-\icmlauthor{Sarah Nadi}{snadi@uwaterloo.ca}
-\icmladdress{University of Waterloo,
-            Waterloo, ON, Canada}
-
-\vskip 0.3in
-]
+\author{Sarah Nadi\\snadi@uwaterloo.ca\\CS886 - Fall 2010\\University of Waterloo\\ Ontario, Canada}        
+\thispagestyle{empty}
+\maketitle
 
 \begin{abstract} 
-STILLLL
+When changing one component in an IT system, an analyst must be careful not to negatively impact any of the other system components. Detecting the complete
+change set (all those components that need to be changed to avoid any side effects) is therefore very important. Given historic information about how different
+components of a system change together, we build a Bayesian network that can be used to predict this change set. Given an initial component to be changed in
+the system, the posterior probability of the other components changing can be calculated using the network. In this paper, we experiment with different
+techniques for building the Bayesian network based on the data available, and report on their predictive power in terms of recall and precision.
 \end{abstract}
@@
 software and hardware are added. In a large system, such changes can cause other parts of the system to malfunction without the ability of the analyst to
-foresee this side effect. Accordingly, we need to ensure that we consider the impact of all changes before they are implemented, and make sure we adjust the
+foresee this side effect. Accordingly, we need to ensure that we consider the impact of changes before they are implemented, and make sure we adjust the
 potentially affected components accordingly while implementing the change. This leads to the process of change set detection which is finding all the
-components of the system that need to be included in your change set where the change set is the set of components that need to be changed to avoid any side
+components of the system that need to be included in your change set. The change set is the set of components that need to be changed to avoid any side
 effects.
 
-Many of the large organizations with large and sophisticated IT systems employ a Configuration Management Database (CMDB) to help them manage these systems.
+Many of the organizations with large and sophisticated IT systems employ a Configuration Management Database (CMDB) to help them manage these systems.
 The CMDB keeps track of all the components in a system, how they are related, as well as all the changes that have occurred to them. Any component in the CMDB
 is called a Configuration Item (CI). In previous work~\cite{nadi2010}, we mined the CMDB repository for historical co-changes, and used the mined correlations
-to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well based on its change
-history with other CIs. We obtained really good results in terms of recall and precision (69.8\% and 88.5\% respectively).
+to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well. We obtained really
+promising results in terms of recall and precision (69.8\% and 88.5\% respectively).
 
 In this paper, we wish to examine the same problem, but from a Bayesian perspective. Instead of looking at pairwise historical co-changes of CI, we would like
-to consider all the changed CIs in the past to deduce the causality relations between them. That is, what is the probability that if A changes, B changes as
-well. In that case, A would be a parent to B in a Bayesian network. Accordingly, we explore the different ways a Bayesian network can be constructed from the
-data we have, and then test the predictions produced by querying this network to examine the obtained recall and precision. We hope that we can obtain better
-results using Bayesian networks since they are not limited to pairwise comparisons.
+to consider all the changed CIs, and learn the relationships between them. That is, given a set of observations about how different CIs change
+together, can we deduce the relationships between them that would allow us to predict future change sets. Accordingly, we explore the different ways a Bayesian
+network can be constructed from the data we have, and then test the predictions produced by querying this network to examine the obtained recall and precision.
+We hope that we can obtain better results using Bayesian networks since they are not limited to pairwise comparisons.
 
-The rest of this paper is organized as follows. BLA BLA BLA
+The rest of this paper is organized as follows. Section~\ref{sec:bg} provides the background information necessary for this work. It explains the nature of
+CMDBs, Bayesian networks, and the tools we used for analyzing Bayesian networks. Section~\ref{sec:constructing} then presents the techniques we used to build
+our Bayesian network. This includes how we process the data, and build the different Bayesian networks accordingly. Section~\ref{sec:setup} details the
+setup we use for our experiments, and Section~\ref{sec:results} provides the results obtained. Section~\ref{sec:future} provides future work that can be done
+to improve the obtained results. Section~\ref{sec:concl} concludes this paper.
 
 \section{Background}
+\label{sec:bg}
 
 \subsection{CMDBs and Change Sets}
-A Configuration Management Database (CMDB)~\cite{itil} is useful in Enterprise IT Management (EITM) since it provides information about the various critical
-components in a
-system including hardware, software, and services provided by the company. It records the configuration of these items, their change history, their incident
-history, as well as the relationships between them. Each item stored in the CMDB is referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows
-an example CMDB to illustrate the concepts of CIs and relationships.
+The repository of information we are using is the Configuration Management Database (CMDB)~\cite{itil}. The CMDB is useful in Enterprise IT Management (EITM)
+since it provides information about the various critical components in a system including hardware, software, and services provided by the company. It records
+the configuration of these items, their change history, their incident history, as well as the relationships between them. Each item stored in the CMDB is
+referred to as a Configuration Item (CI). Figure~\ref{fig:cmdbExample} shows an example CMDB to illustrate the concepts of CIs and relationships.
 
@@
 Management, and in particular, on the problem of change set detection. A \textit{change} is the addition, modification, or removal of anything that could affect
-on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change one CI is the system, other CIs that might
-need to be changed as well must be correctly identified. The set of CIs that will need to be modified for the change to be complete is called a change set.
+on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change a CI is the system, other CIs that might
+need to be changed as well must be correctly identified. We refer to the set of CIs that will need to be modified for the change to be complete ias the
+\textit{change set}.
 
@@
 
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=15cm]{graphics/constructingmodel.png}
+\caption{Building the Bayesian Network}
+\label{fig:process}
+\end{figure*}
+
+Bayesian Networks, first introduced by Pearl~\cite{pearl1988probabilistic} are a type of probabilistic graphical models. Each node in the network
+represents a random
+variable, and nodes are connected through edges which represent probabilistic dependencies~\cite{benbayesian}. A Bayesian network is a Directed
+Acyclic Graph (DAG), and is essentially defined by two aspects: the structure of the graph and the Conditional Probability Tables (CPTs) probabilistically
+relating connected nodes. In our work, a node in the Bayesian network would be a CI which can take values {``true'',``false''} to donate the probability of it
+changing or not. The CPTs would essentially give the probability of one CI changing given its parents' values.
+
+Given a set of data consisting of observations for our variable space, machine learning techniques can be used to estimate the network. If the structure of the
+network is known, and we have full observability (i.e we can observe values for each variable), then learning the CPTs can be achieved through
+Maximum-likelihood estimation (MLE)~\cite{benbayesian}. MLE tries to maximize the log likelihood of the the training data which can simply be done by
+calculating the frequency of occurrence of the values of each variable given its parents. If the structure is known, but we only have partial observability
+(i.e., missing data), then we can use a variation of MLE which is the Expectation Maximization algorithm~\cite{dempster1977maximum}. If the structure is not
+known in
+the first place, then searching through the model space, and calculating network scores for each produced network is necessary. The simplest algorithm for this
+is a Greedy search algorithm~\cite{friedman1996learning}.
+
+While searching for the best network that fits the training data, scores are needed for the produced networks in order to know which one is the best one. There
+are several types of scores that have been proposed in the literature. Two of the commonly used scores are the Minimum Description Length (MDL)
+score~\cite{rissanen1986stochastic} and the BDe score~\cite{heckerman1995learning}. The MDL score is based on finding the length of data required to store a
+network. It tries to balance the complexity of the induced network with the degree of accuracy of which the network represents the frequencies in the training
+data. Therefore, the lower the MDL score of a network, the better it is. The BDe score is proportional to the posterior probability of each network structure
+being scored, given the data. Accordingly, we are searching for the network that maximizes this probability. For large data sets, both metrics score
+structures similarly~\cite{friedman1996learning}.
+
 \subsection{Bayesian Network Tools}
 
-\subsubsection{Weka}
+There are many machine learning and Bayesian analysis tools available. In this work, we use three of these tools which we briefly describe
+below.
 
-WEKA~\cite{hall2009weka} is a toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
-learning problems. It also provides Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
-Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. We will mainly be using WEKA for learning the CPTs in a fixed
-structure network using the SimpleEstimator algorithm~\cite{written2005data} implemented there. This simple estimates the probability values on edges based on
+\subsubsection{WEKA}
+
+WEKA~\cite{hall2009weka} is a Java toolbox for machine learning. It provides different algorithms for classification and clustering as well as other machine
+learning problems. It has Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
+Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. IN terms of learning CPTs, one of the algorithms WEKA implements is
+the SimpleEstimator algorithm~\cite{witten2005data} implemented which we use in our work. This simply estimates the probability values on edges based on
 the frequency values of each variable given its parents in the training set.
@@
 Banjo~\cite{banjotool} is a tool written in Java which infers the structure of a Bayesian network given training data. It has two different search algorithms:
-Greedy and Simulated Annealing. For either of these search algorithms, it has two methods of proposing a new edge. The first is the ``proposeRandomLocalMove''
-which basically proposes a random addition or removal of an edge. The other is ``proposeAllLocalMoves'' which proposes all possible moves, and only keeps the
-best one (?). Banjo uses the BDe metric to compute a network's score.
+Greedy and Simulated Annealing. Banjo uses the BDe metric to compute a network's score, and produces the highest scoring network found. We used the Greedy
+search algorithm in our experiments.
 
@@
 
-\section{Related Work}
-\label{rel-work}
-
-Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. However, their work is on the level of source code changes. They build three
-different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
-co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies, and then the CPTs
-are learnt using the importance sampling algorithm proposed by Changhe and Marek~\cite{yuan2003importance}. The way static dependencies are defined in their
-case is specific to Java. The third one is essentially the first graph, but updated using the historic change information according to the Expectation
-Maximization (EM) algorithm~\cite{dempster1977maximum}. The second was solely based on historic information where the network is build using a greedy structure
- learning algorithm~\cite{friedman1996learning}. They did some preprocessing to their data such as filtering out large changes (with more than 30 elements
-changed at once) since this was probably an insignificant change.
-
-Zhou et al.~\cite{zhou2008} try to answer a slightly different problem. They do not only look at the probability of other elements changing given a specific
-element, they also add features such as authors, change significance levels etc. and try to predict if two elements are co-changes or not accordingly. Thus,
-their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as co-changes or not.
-They use the K2 algorithm proposed by Cooper et. al~\cite{cooper1992bayesian} to estimate the structure of the Bayesian network, and use the SimpleEstimator
-algorithm built in WEKA~\cite{witten2005data}.
+JavaBayes~\cite{javabayestool} is a Java tool for calculating marginal probabilities and expectations in a Bayesian network. In a network, nodes can be set to
+be observed, and then
+the posterior probability of the remaining nodes can be calculated using Bayes theorem. This is the functionality we need here since the CI that will be
+changed will be considered as observed to be ``true'', and then we need to calculate the posterior of the remaining CIs given this observation.
 
 \section{Constructing the Bayesian Network}
-\label{sec:modelsused}
+\label{sec:constructing}
 
-Before constructing the Bayesian Network, we first had to process and prepare the data we have. Then, in order to construct a Bayesian network to use for
+Before constructing the Bayesian Network, we first have to process and prepare the data we have. In order to construct a Bayesian network to use for
 predictions there are two steps involved. First, determining the structure of the actual network, and then estimating the Conditional Probability Tables
@@
 \subsubsection*{Data Size}
-The original data set is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data was infeasible to work with as no tool
-could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to experiment with Bayesian techniques
-for change set prediction,
-it is sufficient to choose a small representative subset of the data. Accordingly, in order to be able to test things properly, we used observations from three
-months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a short period, there were already 2,841 distinct CIs appearing and
-2,229 observations (i.e. change orders). Therefore, we needed to perform further data preprocessing.
+The data set available to us in the sample CMDB we have is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data
+was infeasible to work with as no tool could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to
+experiment with Bayesian techniques for change set prediction, it is sufficient to choose a small representative subset of the data. Accordingly, in order to be
+able to test things properly, we used observations from three months data from January 1, 2008 to March 31, 2008 to build the model. However, even in such a
+short period, there were already 2,841 distinct CIs appearing and 2,229 observations (i.e. change orders). Therefore, we needed to perform further data
+preprocessing.
 
-\begin{figure}[!t]
-\centering
-\includegraphics[width=7cm]{graphics/constructingmodel.png}
-\caption{Building the Bayesian Network}
-\label{fig:process}
-\end{figure}
-
-First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with 12 different change orders in our training
-observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct CIs. Then,
-in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB perspective.
-However, we ignored three common, and not extremely meaningful relations, which are  ``supports'', ``is location for'', and ``backs up''. Adding the related
-parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs. Accordingly, we just kept CIs tht
-appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use throughout our models for fair
-comparison. Additionally, filtering out CIs meant filtering out some of the observations that did not have any CIs satisfying our criteria. This led to us
-having 1,305 observations instead of 2,229 which was a more manegeable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305
-observations (change sets).
-
-FIGURE FOR DATA PROCESSING
+First, we removed all CIs that have changed less than 12 times within this time frame  (i.e were associated with less than 12 different change orders in our
+training observations). There is no particular reason for choosing 12 as a cut off. It simply gave a feasible data set to deal with. This yielded 120 distinct
+CIs. Then, in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB
+perspective. Adding the related parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs.
+Accordingly, we just kept CIs that appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use
+throughout our models for fair comparison. Additionally, filtering out CIs meant filtering out some of the change orders that did not have any CIs satisfying
+our criteria (i.e. all the CIs that changed in them were outside our 170 CI set). This led to us having 1,305 change orders instead of 2,229 which was more
+manageable set. Accordingly, our training data set consisted of 170 variables (CIs), and 1,305 observations (change sets).
 
@@
 change sets are easy to extract and easy to understand. In terms of an observation, CIs were either marked as ``true'' or ``false'' indicating whether they
-have changed or not. Each observation initially consisted of all the in our data set marked as ``false''. For each change order in the training set, we would
+have changed or not. Each observation initially consisted of all the CIs in our data set marked as ``false''. For each change order in the training set, we
+would
 mark each CI appearing in the ``Configuration Items'' field as true, while all those not appearing were left as ``false''. Therefore, the set of training
@@
 \label{sec:structure}
-There were different ways in which we could estimate the structure of the network, and so we built a Bayesian Network using each technique.
+There were essentially three different ways in which we could estimate the structure of the network. We, thus, built a Bayesian Network using each technique to
+be able to compare their performance.
 
-\subsubsection{Model 1 (Banjo)}
+\subsubsection{Method 1 (Banjo)}
 \label{sec:model1}
 
-For the first model, we used Banjo to learn the structure of the Bayesian network from the training data. We use the Greedy searching algorithm implemented
+For the first method, we used Banjo to learn the structure of the Bayesian network from the training data. We use the Greedy searching algorithm implemented
 there. Banjo produced five different top-scoring networks (having the same score), and we simply chose one of them as the network to be used.
 
-\subsubsection{Model 2 (CMDB Relationships)}
+\subsubsection{Method 2 (WEKA Classification)}
 
-For the second model, we used the relations existing in the CMDB to infer the structure of the network (that is place the edges between the nodes). We used two
-variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is if there is an edge from A to B, we will place an
-edge in the Bayesian network from A to B (we will call this model 2A). The second was to reverse the direction of the relationship edges in the CMDB. That is,
-if there is an edge from A to B, we will place an edge in the Bayesian network from B to A (we will call this model 2B). In both cases, however, we chose to
-ignore some relationship edges (WHY, TRY WITHOUT). There was one problem, however, with the generated network. Two of the CIs had more than 15 parents to them
-which means that their CPTs will be intractable to compute. For those two CIs, we simply removed all their parents to make the computation tractable.
+Although our research problem as currently defined is not really a classification problem, we want to examine how the network built by the BayesNet classifier
+in WEKA will perform. Accordingly, we estimated the structure of the network using the K2 structure learning algorithm implemented in WEKA which is also a
+greedy
+search based learning algorithm. The default maximum number of parents allowed for any node is one (a Naive Bayesian Network). However, we used a maximum of
+five to allow the nodes to be related since there is really no class label in our data set. We could not experiment with other techniques in WEKA because they
+do not allow multiple parents.
 
-\subsubsection{Model 3 (Missing Data)}
+\subsubsection{Method 3 (CMDB Relationships)}
 
-This model is the same as the second model, except that we experimented with the missing data feature in Weka. Weka allows the user to specify missing values
-for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
-'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. Therefore, this model also has
-two parts: model 3A and model 3B where the first uses the CMDB relationships in their same direction, and the second uses the reverse direction.
+For the third method, we used the relations existing in the CMDB to infer the structure of the network. We used two
+variations for this. The first was to follow the direction of the relationship edges in the CMDB. That is, if there is a relationship edge in the CMDB from A to
+B, we will place an
+edge in the Bayesian network from A to B. The second was to reverse the direction of the relationship edges in the CMDB. That is,
+if there is a relationship edge from A to B in the CMDB, we will place an edge in the Bayesian network from B to A. There was one problem, however, with
+the generated networks. Two of the CIs had more than 15 parents to them which means that their CPTs will be intractable to compute. For those two CIs, we simply
+removed all their parents to make the computation tractable.
 
@@
 
-However, for all
-models, we used the SimpleEstimator algorithm~\cite{written2005data} built in Weka to calculate the
-CPTs. After setting the data set to be the observations in our training set, Weka
-learned the CPTs for the network produced by Banjo. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference.
+After the structure was determined, we used the SimpleEstimator algorithm~\cite{witten2005data} built in WEKA to calculate the CPTs for all networks. After
+setting the data set to be the observations in our training set (an ARFF file), WEKA learned the CPTs for each of the
+networks above. This complete network was then saved as an BIF XML file to be used by JavaBayes for general inference as will be explained in the next section.
+
+Additionally, we experimented with the missing data feature in WEKA. WEKA allows the user to specify missing values
+for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
+'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. When using Banjo or the CMDB
+relationships to estimate the structure of the network, the missing values were only inserted when learning the CPTs. When using WEKA, the missing values were
+inserted both to learn the network, and to estimate the CPTs.
 
 \section{Experiment Setup}
+\label{sec:setup}
+
+This section explains how we tested the predictions of each of the networks in a way that simulates how an analyst would use the tool in reality. We first
+explain how we simulate this behavior, and then explain how we evaluate the predicted CIs. Algorithm~\ref{alg:simulation} summarizes these steps taken to
+predict a change order.
 
@@
 \begin{algorithmic}
-   \STATE {\bfseries Input:} changeOrder
-    \STATE {\bfseries Input:} BayesianNetwork
-    \STATE {\bfseries Input:} threshold
-    \STATE Set initialCI = first CI in changeOrder
-    \STATE Set occurredSet = changeOrder - initialCI
-    \STATE Initialize observedSet = {firstCI}
-    \STATE Initialize predictedSet = {}
+   \STATE {\bfseries Input:} $changeOrder$
+    \STATE {\bfseries Input:} $BayesianNetwork$
+\STATE{\textbf{Input:}} $networkCIs$ 
+    \STATE {\bfseries Input:} $threshold$
+    \STATE Set $initialCI$ = first CI in changeOrder
+    \STATE Set $occurredset$ = $changeOrder$ - $initialCI$
+    \STATE Initialize $observedSet$ = \{$initialCI$\}
+    \STATE Initialize $predictedSet$ = \{\}
    \REPEAT  
-    \STATE Initialize newPredictions = {}
-    \FOR{$node$ {\bfseries to} $BayesianNetwork$}
-    \STATE posterior = perform inference using Banjo
-    \IF{$posterior > threshold$} 
-      \STATE Add node to newPredictions   
+    \STATE Initialize $newPredictions$ = \{\}
+ \IF{$initialCI$ not in $networkCIs$} 
+\STATE return 
+\ENDIF
+    \FOR{$node$ in $BayesianNetwork$}
+    \STATE $posterior =$ query $node$ in $BayesianNetwork$ 
+    \IF{$posterior > threshold$ \textbf{and} $node$ \textbf{not in} $predicted set$} 
+      \STATE Add $node$ to $newPredictions$   
     \ENDIF
@@
 
-  \FOR{$prediction$ {\bfseries to} $newPredictions$}
-    \IF{$prediction  occurredSet$}
-      \STATE Add prediction to observedSet
+  \FOR{$prediction$ in $newPredictions$}
+    \IF{$prediction$ in  $occurredSet$}
+      \STATE Add $prediction$ to $observedSet$
     \ENDIF
     \ENDFOR
-\STATE Add newPredictions to predictedSet
+\STATE Add $newPredictions$ to $predictedSet$
 
-   \UNTIL{$newPredictions$ is $empty$}
+   \UNTIL{$newPredictions$ is \{\}}
+\newline
+\STATE $recall=  \frac{predictedSet\cap occurredSet}{occurredSet}$
+\newline
+\STATE $precision = \frac{predictedSet\cap occurredSet}{predictedSet} $
 \end{algorithmic}
@@
 accept or reject these CIs (i.e add them to the change set or not). Based on the CIs added to the change set, the analyst can ask the tool for more
-suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''.
+suggestions. This is along the lines of ``Now, that I am also going to change these CIs, what else do I need to change?''. This process is explained in more
+details by Nadi et al.~\cite{nadi2010}.
 
 To simulate this process for our experiments, we do the following. At this point, we have the Bayesian network ready, and we would like to perform Inference.
-More formally, given that a CI will change (our observation), we want to infer the probability that the other CIs in the network might change as well.
-Unfortunately, neither of the two tools previously used provide an Bayesian inference engine. We, therefore, use JavaBayes in this step since it accepts the
-same ARFF format used by Weka. We using a one month testing set where we try to predict all the change orders in April 2008. There was a total of
-883 change orders in that month.
+More formally, given that a CI will change (our observation), we want to update our beliefs that the other CIs in the network might change as well.
+Unfortunately, neither of the two tools previously used provide a general inference engine. We, therefore, use JavaBayes in this step since it accepts
+the same XML BIFF format produced by WEKA. We use a one month testing set where we try to predict all the change orders in April 2008. There was a total of
+883 change orders in that month which we use in our testing set.
 
-For each change order, we would take the first CI as the initial CI to change, then we would set that as an observed node, and update
-the beliefs of all the nodes (CIs) in the network using JavaBayes. We would then loop on all the updated CIs, and add those that match our threshold criteria to
-the predicted change set. To simulate a real life scenario, we then checked which of these predicted CIs actually lies in the target change set we are trying to
-predict. This is similar to an analyst accepting CIs into their change set. These common CIs would then also be marked as observations so that we can predict
-now that I'm going to change these accepted CIs as well, what else do I need to change. Again, we would calculate the posterior probability, and continue doing
-so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst or not) are part of the predicted
-set.
+For each change order, we would take the first CI recorded as the initial CI to change, then we would set that as an observed node, and update
+the beliefs of all the nodes (CIs) in the network using JavaBayes. We are basically checking the posterior probability of each CI being ``true'' given our
+observation. We used different cutoff thresholds for the posterior probabilities we would consider. We would then loop on all the updated CIs, and add those
+that match our threshold criteria to the predicted change set. For example a threshold of 0.2 means that only nodes that have a posterior probability greater
+than 0.2 of having the value ``true'' will be returned. To simulate a real life scenario, we then checked which of these predicted CIs actually lies in the
+target change set we are trying to predict. This is similar to an analyst accepting CIs into their change set. These common CIs would then also be marked as
+observed so that we can predict what else will need to change given that these CIs are also changing. Again, we would calculate the posterior probability given
+the new observations, and continue doing so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst
+or not) are part of the predicted set. These steps (shown in Algorithm~\ref{alg:simulation} would be repeated for each change order. Note that if the initial
+CI is not in the list of CIs we used for training, then the predicted set will be empty since we have not seen this CI before.
 
@@
 
-We need a way to evaluate the predicted CIs. The recall and precision measures from the information retrieval field are appropriate for this type of evaluation.
+We need a way to evaluate the predicted CIs. The recall and precision measures from Information Retrieval are appropriate for this type of evaluation.
 Recall measures the proportion of correct CIs retrieved by the system, while precision measures the proportion of suggested CIs that are correct~\cite{van79}.
 
-Similar to Hassan et. al~\cite{hassan2004predicting}, we define the \textit{Predicted Set} (P) as the set of all CIs DRACA suggests through the full the
-iteration process (see Figure~\ref{fig:process}). We define the \textit{Occurred Set} (O) as the CIs remaining in the change set after excluding the Initial CI
+Similar to Hassan et. al~\cite{hassan2004predicting}, we define the \textit{Predicted Set} (P) as the set of all predicted. We define the \textit{Occurred
+Set} (O) as the CIs remaining in the change set after excluding the Initial CI
 provided by the analyst (i.e Change Set - Initial CI). The intersection of the predicted set and the occurred set, called $PO$, is the common CIs in both sets.
@@
 
-Algorithm~\ref{alg:simulation} shows this process, and how the recall and precision are calculated accordingly.
+\begin{table*}[h!t!]
+\centering
+\resizebox{\textwidth }{!}{
+\begin{tabular}{ c | c |c ||c | c || c |c ||c | c || c |c ||c| c||c|c||c|c}
+\textbf{Threshold} 
+& \multicolumn{2}{|c||}{\textbf{Banjo}} 
+& \multicolumn{2}{|c||}{\textbf{Banjo Missing}}
+&\multicolumn{2}{|c||}{\textbf{WEKA}}
+&\multicolumn{2}{|c||}{\textbf{WEKA Missing}}
+&\multicolumn{2}{|c||}{\textbf{CMDB}}
+&\multicolumn{2}{|c||}{\textbf{CMDB Missing}}
+&\multicolumn{2}{|c||}{\textbf{Reversed CMDB}} 
+&\multicolumn{2}{|c}{\textbf{Reversed CMDB Missing}}\\
+\hline
+ & \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall}
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}
+& \textbf{Recall} 
+& \textbf{Precision}\\
+\hline\hline
+0&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473&0.715&0.473\\
+\hline
+0.1&0.649&0.493&0.715&0.473&0.707&0.501&0.715&0.473&0.569&0.470&0.715&0.473&0.583&0.471&0.715&0.473\\
+\hline
+0.2&0.643&0.501&0.715&0.473&0.704&0.509&0.715&0.473&0.568&0.470&0.715&0.473&0.580&0.470&0.715&0.473\\
+\hline
+0.3&0.634&0.510&0.715&0.473&0.700&0.519&0.715&0.473&0.568&0.469&0.715&0.473&0.569&0.469&0.715&0.473\\
+\hline
+0.4&0.624&0.511&0.715&0.473&0.696&0.524&0.715&0.473&0.568&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.5&0.610&0.500&0.715&0.473&0.682&0.528&0.715&0.473&0.568&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.6&0.607&0.500&0.715&0.473&0.674&0.529&0.715&0.473&0.567&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.7&0.595&0.499&0.715&0.473&0.646&0.526&0.715&0.473&0.567&0.469&0.715&0.473&0.568&0.469&0.715&0.473\\
+\hline
+0.8&0.595&0.499&0.715&0.473&0.627&0.526&0.715&0.473&0.567&0.469&0.715&0.473&0.567&0.469&0.714&0.473\\
+\hline
+0.9&0.574&0.475&0.715&0.473&0.588&0.489&0.567&0.469&0.567&0.469&0.715&0.473&0.567&0.469&0.712&0.473\\
+\hline
+1&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000&0.567&1.000\\
+\end{tabular}
+}
+\centering
+\caption{Summary of Results}
+\label{tab:results}
+\end{table*}
 
@@
 
+\begin{figure}[!t]
+\centering
+\includegraphics[width=8cm]{graphics/cmdbresults.png}
+\caption{F-measure of the different Bayesian networks produced using the CMDB relationships}
+\label{fig:cmdbresults}
+\end{figure}
+
+For all the networks generated from the different methods explained in Section~\ref{sec:structure}, and following the procedure described in
+Section~\ref{sec:setup}, we calculated the average recall and precision for all the change orders in the test set. Based on the average recall and precision
+values, we calculated the F-measure. Table~\ref{tab:results} summarizes the recall and precision obtained for all models. For better visualization, we plot the
+F-measure from all the networks based on CMDB structure together, and the F-measure from those based on Banjo and WEKA together.
+
+Figure~\ref{fig:cmdbresults} shows the F-measure obtained from the four different networks built based on CMDB relationships for structural information. The
+first observation is that reversing the relationship edges did not produce any difference in the results. The curves for the CMDB relationships and the reversed
+CMDB relationships are completely overlapping. The same things applies to the curves of CMDB relationships with missing data, and reversed CMDB relationships
+with missing data. The second observation here is that denoting CIs that did not appear in the change order as missing data rather than giving them the value
+``false'' produced slightly better results. However, the main conclusion from these results is that given each change order, all of the predicted CIs in the
+network already had a posterior probability of greater than 0.9. That is why the F-measure is the same for all thresholds (excluding 0), and only
+changes when the threshold becomes 0.9. This seems rather strange, and despite the relatively high F-measure, it does not seem that the structure
+of these networks is reflective of their relationships in terms of change.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=8cm]{graphics/banjowekares.png}
+\caption{F-measure from network structures estimated by WEKA versus Banjo}
+\label{fig:wekabnj}
+\end{figure}
+
+Figure~\ref{fig:wekabnj} compares the F-measure obtained from the graphs whose structure was built using Banjo versus those built using WEKA. It also shows the
+effects including missing data in both cases. Surprisingly, it seems that the network built using WEKA has better performance (although only slightly better).
+However, marking CIs that have not been changed as unobserved (missing) instead of ``false'' slightly deteriorated the predictive performance. Overall, the
+performance of the two networks are very similar which comes as a a bit of a surprise since WEKA is really a classification engine and assumes that the last
+variable in the data is the target class which is not the case in our data.
+
+When comparing networks built from CMDB structure, and those built by WEKA or Banjo, we still notice that the performance in terms of the F-measure is very
+close. However, WEKA and Banjo had slightly better recall as shown in Table~\ref{tab:results}. Additionally, while running our tests, it took much longer to
+query the graphs built by WEKA and Banjo that those built from the CMDB structure. This implies that the former graphs have more interconnections between the
+nodes while the latter are more sparse.
+
+We also note when the threshold is 1.0 (i.e no predictions are being made), the recall is not 0 although the precision is 1. This is because there are many
+change sets of size 1 which makes the occurred set of size 0, and thus recall being automatically 1. Please refer to Nadi et. al~\cite{nadi2010} for more
+details. Additionally, all networks start at the same recall and precision values when the threshold is 0 since simply all 170 CIs are predicted. This is
+because in the Bayesian network, the probability of any CI is never really zero. It may be very close to zero, but will never be completely unlikely.
+
+From looking at the results, we were a bit worried that most of the predictions were empty sets, because the initial CI was not seen before in the training
+data. Therefore, we thought of limiting the test data to only those nodes that appeared in the training, and adjusting the observations in the test set
+accordingly. In that case, for almost all models, we would either see perfect recall and almost zero precision or perfect precision and the recall obtained
+when no predictions are made. This indicates that limiting the test data to the same set of CIs used in training leads to some sort
+of overfitting. These results are not shown here due to space limitations.
+
+\section{Related Work}
+\label{rel-work}
+
+Mirarab et al.~\cite{mirarab2007} investigate the same problem as our work. They build three
+different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
+co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies, and then the CPTs
+are learnt using the importance sampling algorithm proposed by Changhe and Marek~\cite{yuan2003importance}. The way static dependencies are defined in their
+case is specific to Java. The third one is essentially the first graph, but updated using the historic change information according to the Expectation
+Maximization (EM) algorithm~\cite{dempster1977maximum}. The second was solely based on historic information where the network is build using a greedy structure
+ learning algorithm~\cite{friedman1996learning}. The main difference between their work and ours is that their work is on the level of source
+code while ours is on a system level.
+
+Zhou et al.~\cite{zhou2008} try to answer a slightly different problem. They do not only look at the probability of other elements changing given a specific
+element, they also add features such as authors, change significance levels etc. and try to predict if two elements are co-changes or not accordingly. Thus,
+their problem is more of a classification problem where given two elements, and some observed features they try to determine the class as a co-change or not.
+They use the K2 algorithm proposed by Cooper et. al~\cite{cooper1992bayesian} to estimate the structure of the Bayesian network, and use the SimpleEstimator
+algorithm built in WEKA~\cite{witten2005data}. Although the problems being solved are different, we used similar almost the same techniques they used to build
+the network.
+
+\section{Future Work}
+\label{sec:future}
+
+There is still some work that needs to be done in this area. First, it would be interesting to experiment with undirected graphs. For our problem, when two CIs,
+A and B, occur in the same change order, it is hard to determine whether A caused B to change or B caused A to change. Therefore, using an undirected model,
+such as a Markov network could be interesting in this setting. Unfortunately, due to time constraints, we could not experiment with that. Another possible
+expansion is to use the CMDB relationships more carefully when determining the structure of the network. Currently, we either add edges to the Bayesian network
+following the direction of those in the CMDB or in the reverse direction. It would be interesting to determine which set of relationships semantically make
+sense to have the causality relationship in their same direction, and which make sense to have it in the opposite direction. That way, the structure of the
+network produced could perhaps be more accurate.
+
+Another addition would be to make the model more ``intelligent''. That is, instead of not predicting anything at all for CIs that have not been previously
+seen, it can find a similar CI it has information about, and predict the change set based on it. The challenge here is how to define ``similar''. This can
+simply be the same type of CI (i.e. database server, application, router etc.) with the assumption that CIs of the same type will behave similarly or it can be
+CIs with the same relationships. This would require further investigation as to which criteria is the best for describing similarity here.
+
 \section{Conclusion}
-\label{concl}
+\label{sec:concl}
+
+In this work, we experimented with Bayesian networks in the context of change set predictions for IT systems. Given the Configuration Items (CIs) the analyst
+wants to change in the system, we can predict what other CIs might need to be changed as well. We built different Bayesian networks based on different
+information such as CMDB relationships or using structural learning algorithms such as the Greedy algorithm. Overall, all the networks built gave us, more or
+less, the same results. We believe that this is a good start for this exploratory work which indicates that Bayesian networks can be employed in change set
+detection. However, more work needs to be done in terms of making better predictions for unseen CIs, and in terms of exploring other types of probabilistic
+graphs that might be used in this context. We hope to expand this work to explore more variations of Bayesian learning in order to get better change set
+predictions.
 
 \bibliography{references}
-\bibliographystyle{icml2010}
+\bibliographystyle{ieee}
 

### 2010-12-11 cs886-sarah/Paper.tex
sat 11th after correcting some typos and graph size

@@
 is called a Configuration Item (CI). In previous work~\cite{nadi2010}, we mined the CMDB repository for historical co-changes, and used the mined correlations
-to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well. We obtained really
-promising results in terms of recall and precision (69.8\% and 88.5\% respectively).
+to predict change sets. That is, given an initial CI that is going to change, we predict what other CIs might need to be changed as well. We obtained
+promising results in terms of recall and precision.
 
@@
 our Bayesian network. This includes how we process the data, and build the different Bayesian networks accordingly. Section~\ref{sec:setup} details the
-setup we use for our experiments, and Section~\ref{sec:results} provides the results obtained. Section~\ref{sec:future} provides future work that can be done
+setup we use for our experiments, and Section~\ref{sec:results} provides the results obtained. Section~\ref{sec:future} suggests future work that can be done
 to improve the obtained results. Section~\ref{sec:concl} concludes this paper.
@@
 A CMDB provides a basis for decision making processes such as Incident Management, Change Management, etc.  In this paper, we focus on the process of Change
-Management, and in particular, on the problem of change set detection. A \textit{change} is the addition, modification, or removal of anything that could affect
-on IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change a CI is the system, other CIs that might
+Management, and in particular, on the problem of change set detection. A \textit{change} is the addition, modification, or removal of anything that could
+affect IT services. A poorly planned change may lead to a fault in the system. Accordingly, when one wants to change a CI is the system, other CIs that might
 need to be changed as well must be correctly identified. We refer to the set of CIs that will need to be modified for the change to be complete ias the
@@
 learning problems. It has Bayesian learning techniques for classification problems. Additionally, it provides explorations tools for Bayesian
-Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. IN terms of learning CPTs, one of the algorithms WEKA implements is
-the SimpleEstimator algorithm~\cite{witten2005data} implemented which we use in our work. This simply estimates the probability values on edges based on
+Networks, and allows the user to learn the structure and CPTs of a Bayesian Network. In terms of learning CPTs, one of the algorithms WEKA implements is
+the SimpleEstimator algorithm~\cite{witten2005data} which we use in our work. This simply estimates the probability values on edges based on
 the frequency values of each variable given its parents in the training set.
@@
 \subsubsection*{Data Size}
-The data set available to us in the sample CMDB we have is from three years, and has 7,999 distinct CIs, and 27,305 change orders. This amount of data
+The data set available to us in the sample CMDB we have is from three years (Jan 2006 - Dec. 2008), and has 7,999 distinct CIs, and 27,305 change orders.
+This amount of data
 was infeasible to work with as no tool could handle such a large amount of variables in a Bayesian network. For the purposes of this project, which is mainly to
@@
 CIs. Then, in order to slightly increase our variable space to include other related CIs, we found all the parent CIs related to these 120 CIs from the CMDB
-perspective. Adding the related parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may not be in the original set of 2,841 CIs.
+perspective (i.e. had a relationship edge in the CMDB). Adding the related parent CIs, we now had a set of 241 CIs. However, some of the added parent CIs may
+not be in the original set of 2,841 CIs.
 Accordingly, we just kept CIs that appeared more than 12 times or were in the set of related CIs. This provided us our final data set of 170 CIs which we use
@@
 \label{sec:structure}
-There were essentially three different ways in which we could estimate the structure of the network. We, thus, built a Bayesian Network using each technique to
+There were essentially three different ways in which we could estimate the structure of the network. We, thus, built a Bayesian network using each technique to
 be able to compare their performance.
@@
 
-Although our research problem as currently defined is not really a classification problem, we want to examine how the network built by the BayesNet classifier
+Although our research problem as currently defined is not really a classification problem, we wanted to examine how the network built by the BayesNet classifier
 in WEKA will perform. Accordingly, we estimated the structure of the network using the K2 structure learning algorithm implemented in WEKA which is also a
@@
 search based learning algorithm. The default maximum number of parents allowed for any node is one (a Naive Bayesian Network). However, we used a maximum of
-five to allow the nodes to be related since there is really no class label in our data set. We could not experiment with other techniques in WEKA because they
+five to allow the nodes to be related since there is really no class label in our data set. We could not experiment with other structural learning techniques in
+WEKA because they
 do not allow multiple parents.
@@
 if there is a relationship edge from A to B in the CMDB, we will place an edge in the Bayesian network from B to A. There was one problem, however, with
-the generated networks. Two of the CIs had more than 15 parents to them which means that their CPTs will be intractable to compute. For those two CIs, we simply
+the generated networks. Two of the CIs had more than 15 parents to them which made their CPTs more intractable to compute. For those two CIs, we simply
 removed all their parents to make the computation tractable.
@@
 Additionally, we experimented with the missing data feature in WEKA. WEKA allows the user to specify missing values
-for any variable by simply putting a '?' instead of its value. Therefore, instead of putting the CIs that did not appear in a change order as false, we put a
-'?' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. When using Banjo or the CMDB
+for any variable by simply putting a '?' instead of its value. Therefore, instead of marking the CIs that did not appear in a change order as ``false'', we put
+a
+``?'' in their place. This seemed closer to practice, because in reality, we are not sure whether this CI actually changed or not. When using Banjo or the CMDB
 relationships to estimate the structure of the network, the missing values were only inserted when learning the CPTs. When using WEKA, the missing values were
@@
 the same XML BIFF format produced by WEKA. We use a one month testing set where we try to predict all the change orders in April 2008. There was a total of
-883 change orders in that month which we use in our testing set.
+883 change orders in that month which we use in our testing set. No preprocessing was performed on the testing data.
 
@@
 observed so that we can predict what else will need to change given that these CIs are also changing. Again, we would calculate the posterior probability given
-the new observations, and continue doing so until there are no more common CIs. All the CIs that match the threshold criteria (whether accepted by the analyst
-or not) are part of the predicted set. These steps (shown in Algorithm~\ref{alg:simulation} would be repeated for each change order. Note that if the initial
+the new observations, and continue doing so until there are no more new CIs to predict. All the CIs that match the threshold criteria (whether accepted by the
+analyst
+or not) are part of the predicted set. These steps (shown in Algorithm~\ref{alg:simulation}) would be repeated for each change order. Note that if the initial
 CI is not in the list of CIs we used for training, then the predicted set will be empty since we have not seen this CI before.
@@
 Section~\ref{sec:setup}, we calculated the average recall and precision for all the change orders in the test set. Based on the average recall and precision
-values, we calculated the F-measure. Table~\ref{tab:results} summarizes the recall and precision obtained for all models. For better visualization, we plot the
+values, we calculated the F-measure. Table~\ref{tab:results} summarizes the recall and precision obtained for all models rounded to three decimal places. For
+better visualization, we plot the
 F-measure from all the networks based on CMDB structure together, and the F-measure from those based on Banjo and WEKA together.
@@
 
-From looking at the results, we were a bit worried that most of the predictions were empty sets, because the initial CI was not seen before in the training
-data. Therefore, we thought of limiting the test data to only those nodes that appeared in the training, and adjusting the observations in the test set
+Given that there will be many CIs in the testing data that have not been seen before in the training, we thought of limiting the test data to only
+those nodes that appeared in the training, and adjusting the observations in the test set
 accordingly. In that case, for almost all models, we would either see perfect recall and almost zero precision or perfect precision and the recall obtained
@@
 different Bayesian Networks, one that is based on package and class dependency information (static relationships), one which is dependent on historical
-co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies, and then the CPTs
+co-changes, and one which uses both. For the first graph, the initial structure is essentially ``given'' according to the static dependencies in the code, and
+then the CPTs
 are learnt using the importance sampling algorithm proposed by Changhe and Marek~\cite{yuan2003importance}. The way static dependencies are defined in their
@@
 
-There is still some work that needs to be done in this area. First, it would be interesting to experiment with undirected graphs. For our problem, when two CIs,
+There is still a lot of work that needs to be done in this area. First, it would be interesting to experiment with undirected graphs. For our problem, when two
+CIs,
 A and B, occur in the same change order, it is hard to determine whether A caused B to change or B caused A to change. Therefore, using an undirected model,
@@
 simply be the same type of CI (i.e. database server, application, router etc.) with the assumption that CIs of the same type will behave similarly or it can be
-CIs with the same relationships. This would require further investigation as to which criteria is the best for describing similarity here.
+CIs with the same relationships. This would require further investigation as to which criteria is the best for describing similarity in this context.
 
@@
 information such as CMDB relationships or using structural learning algorithms such as the Greedy algorithm. Overall, all the networks built gave us, more or
-less, the same results. We believe that this is a good start for this exploratory work which indicates that Bayesian networks can be employed in change set
+less, the same results (an F-measure of about 0.6). We believe that this is a good start for this exploratory work which indicates that Bayesian networks can be
+employed in change set
 detection. However, more work needs to be done in terms of making better predictions for unseen CIs, and in terms of exploring other types of probabilistic
